---
title: "Initial_Cleaning-NS-IBTS_data"
output: html_document
editor_options: 
  chunk_output_type: inline
---

#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))


knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))

```

#library
```{r}

library("tidyverse")

```

#------------Read & Merge data--------------------
# ns swc ebs data
(suppressed:bal)
```{r}
#------- NS
ns <-read.csv("data/raw_data/ns_raw_formatted.csv") # 1822998      13
ns <- ns %>% select(-X)
names(ns)

#------- SWC
swc <- read.csv("data/raw_data/swc_raw_formatted.csv")
swc <- swc %>% select(-X)
names(swc)

#------- Bal
bal <- read.csv(file="data/raw_data/bal_raw_formatted.csv")
bal <- bal %>% select(-X)
names(bal)

#------- EBS
ebs <- read.csv(file="data/raw_data/ebs_raw_formatted.csv")
ebs<- ebs %>% select(-X) %>%
  mutate_at("subarea", as.character)
names(ebs)


ns %>% group_by(species) %>% tally()
swc %>% group_by(species) %>% tally()
bal %>% group_by(species) %>% tally()
ebs %>% group_by(species) %>% tally()

```
# Bind all raw data
```{r}
# bind all datasets
fish.all <- bind_rows(ns, swc, bal, ebs)
#fish.all <- bind_rows(ns, swc, ebs)
```

#------------Cleaning Start--------------------

# Select year, quarter
@ Strange for swc, some sp. have same max and min length?
*Because they are too rare*
```{r}

fish.all <-  fish.all %>%
  filter(year >=1982 & year <= 2019) %>% 
  filter(length > 0) %>% 
  filter(cpue_number_per_hour > 0) %>%
  # quarter 1 for ns and swc, quarter 0 for BS and EBS
  filter(quarter == c("1","0")) %>%
  mutate(yearqua=paste(year,quarter,sep=".")) %>%
  relocate(yearqua,.before = "subarea") %>%
  #phylo.ordered(ns.sp.phylo.order) %>%
  mutate_at("species", as.character)

# Check max & min length
fish.all %>%
  group_by(species, region) %>%
  summarise(max_length = max(length),
            min_length = min(length)) %>% 
  arrange(region)
```

# Deriving extreme length & cut-offs

Lower & Upper cut-offs of extreme body size are defined as  lenth shorter than x% or longer than (100-x)% within yearly length-abundance relationship

step1. Deriving yearly cummulative length-abundance relationship
step2. Recording length in 0.1%,1%,3% at both ends in each year as lower/upper bound of extreme length
Step3. Taking median on yearly lower/upper bound (length_upperbound_year; length_lowerbound_year)
step4. Recording Max/Min length across survey records

```{r}

fish.all <- fish.all %>%
  #Step1
  group_by(region, species,year,length) %>% 
  mutate(cpue_length_year=sum(cpue_number_per_hour)) %>%
  arrange(region, species,year, length) %>% 
  #Step2
  group_by(region, species,year) %>% 
  mutate(cummulative_cpue_length = cumsum(cpue_length_year)/sum(cpue_length_year))%>%
  group_by(region, species,year) %>%
  mutate(
         length_lowerbound_year_0.1= length[min(which(cummulative_cpue_length>0.001))],
         length_upperbound_year_0.1= length[min(which(cummulative_cpue_length>0.999))],
         
         length_lowerbound_year_1= length[min(which(cummulative_cpue_length>0.01))],
         length_upperbound_year_1= length[min(which(cummulative_cpue_length>0.99))],
         
         length_lowerbound_year_3= length[min(which(cummulative_cpue_length>0.03))],
         length_upperbound_year_3= length[min(which(cummulative_cpue_length>0.97))]
         
         #length_lowerbound_year_5= length[min(which(cummulative_cpue_length>0.05))],
         #length_upperbound_year_5= length[min(which(cummulative_cpue_length>0.95))],
         
         #length_lowerbound_year_10= length[min(which(cummulative_cpue_length>0.1))],
         #length_upperbound_year_10= length[min(which(cummulative_cpue_length>0.9))]
         
         
         ) %>%
  #Step3
  group_by(region, species) %>% 
  mutate(
         median_length_lowerbound_0.1=median(length_lowerbound_year_0.1),
         median_length_upperbound_0.1=median(length_upperbound_year_0.1),
         
         median_length_lowerbound_1=median(length_lowerbound_year_1),
         median_length_upperbound_1=median(length_upperbound_year_1),
         
         median_length_lowerbound_3=median(length_lowerbound_year_3),
         median_length_upperbound_3=median(length_upperbound_year_3)
         
         #median_length_lowerbound_5=median(length_lowerbound_year_5),
         #median_length_upperbound_5=median(length_upperbound_year_5),
         
         #median_length_lowerbound_10=median(length_lowerbound_year_10),
         #median_length_upperbound_10=median(length_upperbound_year_10)
         
         ) %>%
  #Step4
  group_by(region, species) %>%
  mutate(min_length = min(length),
         max_length = max(length)) 

#fish.all %>% 
#  select(region, species,
#         median_length_lowerbound,
#         median_length_upperbound,
#         min_length,
#         max_length) %>%
#  group_by(region, species) %>% 
#  unique()

```

#Derive extreme.length.cutoff
```{r}
extreme.length.cutoff <- fish.all %>%
  select(region, species, 
         median_length_lowerbound_0.1, median_length_upperbound_0.1,
         median_length_lowerbound_1, median_length_upperbound_1,
         median_length_lowerbound_3, median_length_upperbound_3
         #median_length_lowerbound_5, median_length_upperbound_5,
         #median_length_lowerbound_10, median_length_upperbound_10
         ) %>%
  unique() 
```

#Save extreme.length.cutoff
```{r}

write.csv(extreme.length.cutoff,file= "data/initial_data/extreme_length_cutoff.csv")
```


#Derive (max) range size each year:
step1. Count yearly range for each sp
step2. Derive max yearly range for each sp

This dataset will be used to draw abundance-length distribution
*Note that some species have less than 10 length groups in some years, better to partition them into 5 lengthgroups

```{r}
fish.all <- fish.all %>% 
  #Step1
  group_by(region, year,quarter,species) %>%
  mutate(yearly_range = n_distinct(subarea)) %>%
  #Step2
  group_by(region, species) %>% 
  mutate(max_yearly_range = max(yearly_range))
```

#Save (max) range size each year
```{r}
write.csv(fish.all,file= "data/initial_data/all_region_initial_clean.csv")
```


#------------Below are skippable---------
# Make a summary table
```{r} 

(table <- fish.all %>%
    group_by(region, species) %>%
    summarise(# Number of subareas
              number_subarea_found = n_distinct(subarea),
              # Time series length
              number_year_found = n_distinct(yearqua),
              record_year_start = min(year),
              record_year_end = max(year)))

# save as "species_list_before_filtering_rmd1" in summary_data folder
write.csv(table,file= "data/summary_data/species_list_before_filtering_rmd1.csv")

```

# Number of species per region
```{r}
(table <- fish.all %>%
    group_by(region) %>%
    summarise(number_subarea_found = n_distinct(species)))

# save as "number_species_before_filtering_rmd1" in summary_data folder
write.csv(table,file= "data/summary_data/number_species_before_filtering_rmd1.csv")


```













