---
title: "02Edit_LengthGroupCutting"
output: html_document
editor_options: 
  chunk_output_type: inline
---
#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))


knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))
```

#library
```{r}

library("tidyverse")


```

#------------Functions--------------------
# <Function>: Cut population into N LengthGroup, "Including extreme.length"
step1. Split original dataset into Extreme/Normal Length
step2. For Extreme Length, classified those shorter than left cut-off or longer than right cut-off into LengthGroup "1" or "N"
step3. For Normal Length, classified individuals into LengthGroup "1"~"N"
step4. Combine Extreme/Normal Length data

#Calculate Quotient = ((original length class - min class) %/% (length range/group.num)) +1
```{r}

cut_length_with_extreme <- function(fish.data,group.num){

#Step1
extreme.length <- fish.data %>%
  group_by(species) %>%
  filter(length < median_length_lowerbound | length > median_length_upperbound)#

normal.length <- fish.data %>%
  group_by(species) %>%
  filter(length >= median_length_lowerbound & length <= median_length_upperbound)

#Step2
cpue.length.subarea.yearqua.extreme <- extreme.length %>% 
  group_by(region,species) %>%
  mutate(length_group = case_when(
    length <= median_length_lowerbound ~ 1,
    length >= median_length_upperbound ~ group.num,
  ))

#Step3
cpue.length.subarea.yearqua.no.extreme <- normal.length %>% 
  group_by(region,species) %>%
  mutate(length_group=((length-min(length)) %/% (( (max(length)+0.000001)-min(length))/group.num))+1) %>%
  filter (length_group <= group.num)

#Step4
  cpue.length.subarea.yearqua <- bind_rows(cpue.length.subarea.yearqua.no.extreme,
                                           cpue.length.subarea.yearqua.extreme) %>%
  # Summation of cpue for all original lengthes that belong to the newly-defined Length interval
  group_by(region,species,yearqua,subarea,length_group) %>% 
  summarize(cpue_length_subarea_yearqua=sum(cpue_number_per_hour)) %>%
  
  #phylo.ordered(ns.sp.phylo.order) %>% 
    
  print(paste0("Scenario LengthGroup = ",group.num, " has been completed."))

return(cpue.length.subarea.yearqua)

}


```

#<Function>: Cut population into N LengthGroup, "Excluding extreme.length"
```{r}

cut_length_no_extreme <- function(fish.data,group.num){

normal.length <- fish.data %>%
  group_by(species) %>%
  filter(length >= median_length_lowerbound & length <= median_length_upperbound)#dim(1014561,18)



cpue.length.subarea.yearqua<- normal.length %>% 
  group_by(region,species) %>%
  #Calculate Quotient = ((Original Length Class - min class) %/% (length range/group.num)) +1
  mutate(length_group=((length-min(length)) %/% (( (max(length)+0.000001)-min(length))/group.num))+1) %>% 
  filter (length_group <= group.num) %>%
  # Summation of cpue for all original lengthes that belong to the newly-defined Length interval
  group_by(region,species,yearqua,subarea,length_group) %>% 
  summarize(cpue_length_subarea_yearqua=sum(cpue_number_per_hour)) %>%
  
  print(paste0("Scenario LengthGroup = ",group.num, " has been completed."))

return(cpue.length.subarea.yearqua)

}
#CutInto.LengthGroup.noEx(NS.fish.arrange,10) #dim(533768,6)
```

#------------LengthGroup Cutting--------------------
# Load data
*Input data: all.region.initial_clean.csv*
```{r}
fish <- read.csv("data/initial_data/all_region_initial_clean.csv")

fish <- fish %>%
  select(-X)
```


#Setting different cutoff criteria (for later sensitivity analysis)
```{r}

cutoff <- c(0.1,1,3)

lowerbound <- c(paste0("median_length_lowerbound_",cutoff))
upperbound <- c(paste0("median_length_upperbound_",cutoff))

fish.lower <- fish %>%
  select(region,species,yearqua,subarea,length,
         cpue_number_per_hour, cpue_length_year,
         median_length_lowerbound_0.1,
         median_length_lowerbound_1,
         median_length_lowerbound_3
         #median_length_lowerbound_5,
         #median_length_lowerbound_10
         ) %>%
  gather(key="median_lowerbound_position",
         value = "median_length_lowerbound",
          c(median_length_lowerbound_0.1,
            median_length_lowerbound_1,
            median_length_lowerbound_3
            #median_length_lowerbound_5,
            #median_length_lowerbound_10
            ),
         
         c(-region,-species,-yearqua,-subarea,-length,
           -cpue_number_per_hour,-cpue_length_year)) 


fish.upper <- fish %>%
  select(region,species,yearqua,subarea,length,
         cpue_number_per_hour, cpue_length_year,
         median_length_upperbound_0.1,
         median_length_upperbound_1,
         median_length_upperbound_3
         #median_length_upperbound_5,
         #median_length_upperbound_10
         ) %>%
  gather(key="median_upperbound_position",
         value = "median_length_upperbound",
         c(median_length_upperbound_0.1,
           median_length_upperbound_1,
           median_length_upperbound_3
           #median_length_upperbound_5,
           #median_length_upperbound_10
           ),
         
         c(-region,-species,-yearqua,-subarea,-length,
           -cpue_number_per_hour,-cpue_length_year))


fish <- left_join(fish.lower,fish.upper)


```




# <with loop>: Cut populations into 5 ~ 10 LengthGroups 
```{r}
group.num <- seq(5,10,1)


for ( i in 1:length(group.num)) {
  
  for (j in 1:length(cutoff)) {
  
  fish.data <- fish %>%
    filter(median_lowerbound_position == lowerbound[j] &
           median_upperbound_position == upperbound[j])
  
  fish.cut.length <- cut_length_with_extreme(fish.data = fish.data,
                                           group.num = group.num[i])

  print(paste0("Cut populations into ",group.num[i]," length-groups, row numbers = ",nrow(fish.cut.length[1])))

  dir <- paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_cut.csv")

  write.csv(fish.cut.length,file = dir)
  }
}

```
# <without loop>: Run with 10 groups
```{r}
fish.cut.length <- cut_length_with_extreme(fish.data = fish,
                                       group.num = 5)
# Check
fish.cut.length %>% 
  group_by(region, species) %>% 
  tally()
```





#------------Below are skippable---------
*Note* 
Further cleaning and filtering of "species" & "Years with Incomplete LengthGroup" will be processed in script 03 & 04:
<Process1.>The species will be selected based on mean yearly cpue & yearly summed cpue (script03)
<Process2.>The noise about LengthGroup will be removed based on YearlyNumLengthGroup (script04; also see below Check3.)


*Note* 
<Check1.> ~ <Check3.> are skippable. (Only for familiarizing dataset)
Please see script 3 for further cleaning of fish.cut.length data




# <Check1.> LengthGroup Conditions 
```{r}

fish.cut.length %>% 
  select(region,species,length_group) %>%
  group_by(region,species) %>%
  #Calculate the number of LengthGroup that one species can be partitioned into.
  summarise(num_length_group = n_distinct(length_group)) %>%
  filter(num_length_group<10)

fish.cut.length %>%  
  select(region,species,length_group) %>%
  group_by(region,species) %>%
  summarise(num_length_group= n_distinct(length_group)) %>%
  filter(num_length_group==10)
```


# <Check2.> How many years are each LengthGroup being surveyed
```{r}

fish.cut.length %>% 
  ungroup() %>%select(-subarea,-cpue_length_subarea_yearqua) %>%
  #Count the number of year that one lengthgroup was surveyed across studied period 
  group_by(region,species, length_group) %>% 
  summarise(num_survey_year=n_distinct(yearqua))

  
```






