---
title: "03Edit_SpeciesFiltration_LengthGroupCleaning"
output: html_document
editor_options: 
  chunk_output_type: inline
---

This script filters populations out based on criteria 1-3 (see Methods section of the paper).

#Set working directory
```{r setup}

knitr::opts_knit$set(root.dir = ("/Users/hsiaohang.tao/Dropbox/academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/"))

```

#library
```{r}
library("tidyverse")
```

# species in which the survey data is not reliable 

we removed additional 19 species from the North Sea and the Scottish West Coast due to concerns about the consistency of long-term sampling and reporting in the DATRAS database (Barbara Bland, personal communication, 

```{r}

mich.sp <- c("Leucoraja naevus",
      "Enchelyopus cimbrius",
      "Agonus cataphractus",
      "Scyliorhinus canicula",
      "Echiichthys vipera",
      "Cyclopterus lumpus",
      "Arnoglossus laterna",
      "Sebastes viviparus",
      "Myoxocephalus scorpius",
      "Maurolicus muelleri",
      "Gadiculus argenteus",
      "Lumpenus lampretaeformis",
      "Buglossidium luteum",
      "Callionymus maculatus",
      "Raja montagui",
      "Amblyraja radiata",
      "Raja clavata",
      "Gasterosteus aculeatus",
      "Lycodes vahlii")

```

#------------Functions--------------------
#<Function>: Selection criteria 1: abundance

critical_mean_cpue_year > a

a: the critical abundance value that we could freely define

step1.Calculate total_cpue_year
step2.Calculate mean_cpue_year
step3.Filter species whose mean_cpue_year is above the critical_mean_cpue_year that we defined. 
```{r}
get_key1_mean_cpue_year <- function(fish.data, critical.mean.cpue.year){
 
fish.filter.mean.cpue.year <- fish.data %>%
  #Step1.
  group_by(region, yearqua,species) %>% 
  summarise(total_cpue_year=sum(cpue_length_subarea_yearqua)) %>%
  #Step2.
  group_by(region, species) %>% 
  summarise(mean_cpue_year = mean(total_cpue_year)) %>%
  #Step3.
  filter(mean_cpue_year >= critical.mean.cpue.year) 

  return(fish.filter.mean.cpue.year)
  
}

```


#<Function>: Selection criteria 2: cpue evenness

the number of year reaching "critical.total.cpue.year b" > c

*b: thecritical.total.cpue.year that we could freely define;
*c: LowerBound_NumYear which reaches critical.total.cpue.year

step1.Calculate total_cpue_year
step2.Count num_year during which total_cpue_year reaches the LowerBound_NumYear we defined.
step3.Filter species whose NumYear meeting critical cpue is above the 80% of survey duration. 

*1221 edit: LowerBound_NumYear = 80% x n_distinct(yearqua)* -> 80% yearqua should reach critical.total.cpue.year

```{r}
get_key2_evenness_cpue <- function(fish.data, critical.total.cpue.year){

fish.filter.evenness.cpue <- fish.data %>% 
  #Step1.
  group_by(region, yearqua, species) %>% 
  summarise(total_cpue_year=sum(cpue_length_subarea_yearqua)) %>%
  #Step2.
  group_by(region,species) %>% 
  mutate(num_year_cpue_above_critical = length(which(total_cpue_year>=critical.total.cpue.year)))%>%  
  #Step3.
  group_by(region) %>%
  filter(num_year_cpue_above_critical >= n_distinct(yearqua)*0.8) %>%
  select(region,species,num_year_cpue_above_critical) %>%
  unique()

return(fish.filter.evenness.cpue)
}

```

#<Function>: Species selection criteria 3 : length group number

critical.group.num d; all.group.num e;

*d: Critical LengthGroup Number; 
*e: All LengthGroup Number

step1. Count num_group_year for each species
step2. Count num_year_group_above_critical & num_year_all_group for each species, then arrange data
step3. Filter species with their critical.group.num/all.group.num meeting all survey year/80% survey year
```{r}
get_key3_group_num <- function(fish.data,critical.group.num, all.group.num){
  
fish.filter.group.num <- fish.data %>% 
  #Step1.
  ungroup() %>% 
  select(-subarea,-cpue_length_subarea_yearqua) %>%
  group_by(region,species,yearqua) %>%
  summarise(num_group_year=n_distinct(length_group))%>%
  #Step2.
  group_by(region,species) %>%
  mutate(num_year_group_above_critical=length(which(num_group_year >= critical.group.num)),
         num_year_all_group=length(which(num_group_year == all.group.num))) %>%
  select(-num_group_year) %>%
  unique()

#Step3.
fish.filter.group.num <- fish.filter.group.num %>% 
  group_by(region) %>%
  filter( num_year_group_above_critical == n_distinct(yearqua) & 
          num_year_all_group >= n_distinct(yearqua)*0.8) %>%
  select(-yearqua) %>%
  unique()

return(fish.filter.group.num)

}

```

#------------Species filtration--------------------
Partitioning populations into fewer length-groups can help to include as many as populations.
So, the species list for our analysis will be produced based on 5 length-groups scenario.

However, we will compare the effect of length-group numbers (N=5~10) with sensitivity analysis.

For all length-group Scenario (N=5~10), we will further clean data for these species in Script 4, and proceed deriving overlap indices in script 5.

#<with loop>: Species selection (N=5~10 & cutoff=0.1~3)
step1. read data
step2. species selection via 3 criteria:
        *consider the above 3 selection keys to select species*
        <Key1>: Select species based on mean abundance during survey period
        <Key2>: Select species based on cpue evenness during survey period
        <key3.> Select species based on number of length group each year
step3. save data
```{r}

group.num=seq(5,10,1)
cutoff= c(0.1,1,3)


for (i in 1:length(group.num)) {
  
  for (j in 1:length(cutoff)) {
  
    #step1
    all.fish.length <- read.csv(paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                                       "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_cut.csv"))

    all.fish.length <- all.fish.length %>% select(-X) %>%
      filter(!species %in% mich.sp)


    #step2
    #<Key1>
    fish.selection.mean.cpue.year <- get_key1_mean_cpue_year(all.fish.length,
                                                             critical.mean.cpue.year = 50)
    #<Key2>
    fish.selection.evenness.cpue <- get_key2_evenness_cpue(all.fish.length,
                                                           critical.total.cpue.year = 25)
    #<Key3>
    fish.selection.group.num <- get_key3_group_num(all.fish.length,
                                                   critical.group.num = round(0.8*group.num[i]),
                                                   all.group.num = group.num[i])


    sp.filter.key.temp <- inner_join(fish.selection.mean.cpue.year,fish.selection.evenness.cpue)
    sp.filter.key <- inner_join(sp.filter.key.temp,fish.selection.group.num)

    sp.filter.key <- sp.filter.key %>% relocate("region", .before = "species")


    #step3.
    write.csv(sp.filter.key, file = paste0("data/initial_data/lengthgroup_specific/",
                                           group.num[i],"_lengthgroup/with_extreme/cutoff_",
                                           cutoff[j],"/filtered_species_",group.num[i],"_lengthgroup.csv"))

    }
}

```

