---
title: "04_Nlengthgroup_Cleaning"
output: html_document
---
#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))

knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))
```

#library
```{r}
library("tidyverse")
library("readxl")
```

#------------Functions--------------------
#<Function>: Reduce the noises

Removing Year with number of lengthgroup < X 
(X depends on Num of lengthgroup)

*Require importing critical.group.num vector (depends on length-group Scenario, can be freely defined)*
Step1. Count the number of length-groups in each year (year.denoise.group)
Step2. Count numyear that acheive critical_lengthgroup or all_lengthgroup
Step3. Merge & Remove Year with number of lengthgroup < X 
```{r}

denoise_group_data <- function(fish.data, group.num, critical.group.num){
  #Step1.
  year.denoise.group <- fish.data %>% 
  ungroup() %>% select(-subarea,-cpue_length_subarea_yearqua) %>%

  group_by(region,species,yearqua) %>%
  summarise(num_group_year=n_distinct(length_group)) %>%
    #Step2.
  group_by(region,species) %>%
  mutate(num_year_group_above_critical=length(which(num_group_year >= critical.group.num)),
         num_year_all_group=length(which(num_group_year==group.num))) %>%
  unique()
  #Step3.
  cpue.length.subarea.yearqua <- left_join(fish.data,year.denoise.group)
  
  cpue.length.subarea.yearqua <- cpue.length.subarea.yearqua %>%
     filter(num_group_year >= critical.group.num)
  
  
  print(paste0("For lengthgroup = ",group.num," senario, subset rows from ", dim(fish.data)[1], " to ",
               dim(cpue.length.subarea.yearqua)[1]))
  
  return(cpue.length.subarea.yearqua)
}

```

#------------Remove noises--------------------
#<With Loop> Remove noises of incomplete lengthgroup
*critical.group.num can be freely defined as we wish*

@HHT 2022.03.23------------------------------------
Have error message when running this chunk:
! Can't join on `x$region` x `y$region` because of incompatible types.
ℹ `x$region` is of type <character>>.
ℹ `y$region` is of type <logical>>.

But no error message when running the next chunk <Without Loop> -----------------------------------------------
```{r}
group.num = seq(5,10,1)
cutoff = c(0.1,1,3)

for (i in 1:length(group.num)) {
  for (j in 1:length(cutoff)) {
    
  
    import.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                  "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_cut.csv") 
    all.region.data <- read.csv(import.dir)
    
    import.dir2 = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                  "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/filtered_species_",group.num[i],"_lengthgroup.csv")
    selected.sp <- read.csv(import.dir2) %>% select(region,species)

    all.region.data <- all.region.data %>% inner_join(selected.sp)
  
    all.region.data <- denoise_group_data(all.region.data,
                                          group.num = group.num[i],
                                          critical.group.num = round(0.8*group.num[i]))
  
  
  
    export.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                 "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_denoise.csv")
  
    write.csv(all.region.data, file = export.dir )
  }
}

```

#<Without Loop> Remove noises of incomplete number of lengthgroup
*Run above function step by step*
Step1. Load (1) cpue data with 5 lengthgroup (2) species list
Step2. Filter species with list
Step3. Remove noises & Save data

(demonstrated here with 5 length-groups & 1% cutoff)
```{r}

#Step1.
fish.data <- read.csv("data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_cut.csv")
selected.sp <-  read.csv("data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/filtered_species_5_lengthgroup.csv") %>% select(-X) %>%
  select(region,species)

fish.data <- fish.data %>%
  select(-X) %>% 
  mutate_at("species",as.factor)

#Step2. 
fish.data <- fish.data %>% inner_join(selected.sp)


#Step3.
group.num=5
critical.group.num = 4

cpue.length.subarea.yearqua <- denoise_group_data(fish.data, group.num = 5,
                                                  critical.group.num = 0.8*5)

write.csv(cpue.length.subarea.yearqua,file = "data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_denoise.csv")

```


#------------Below are skippable---------
#Plotting Year-lengthgroup relationship

*1225 edit*
*Check effects of Num_lengthgroup on data reliability*
Reliability:
(1) RangeSize dynamics with year
(2) CPUE_Length dynamics with year
(3) NumYear_AboveCritical
(4) NumYear_AllLength
```{r}

length.group <- list()
group.num = seq(5,10,1)
cutoff = c(0.1,1,3,5,10)
time=1


for ( i in 1:length(group.num)) {
  for (j in 1:length(cutoff)) {
  
  length.group[[time]] <- 
    read.csv(paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                    "_lengthgroup/with_extreme/cutoff_",cutoff[j],
                    "/",group.num[i],"_lengthgroup_denoise.csv")) %>% 
    mutate(num_group=group.num[i],
           cutoff=cutoff[j]) %>%
    relocate(num_group,.before = "yearqua")
  
  time=time+1
  }
  
}


length.group.summary <- bind_rows(length.group) %>%
  filter(cutoff == 10) %>%
  select(-X,-X.1) %>%
  group_by(region,species,num_group,yearqua,length_group) %>% 
  mutate(range_group_year=n_distinct(subarea),
         cpue_group_year=sum(cpue_length_subarea_yearqua)) %>%
  select(-subarea,-cpue_length_subarea_yearqua,-num_group_year) %>%
  unique() %>% 
  mutate(region_abbr=case_when(
    region == "Baltic Sea" ~ "BS",
    region == "Eastern Bering Sea" ~ "EBS",
    region == "North Sea" ~ "NS",
    region == "Scottish West Coast" ~ "SWC"
  )) %>%
  mutate(population=paste0(region_abbr,"-",species)) %>%
  mutate_at("length_group",as.character)


regional.pop.list <- unique(length.group.summary$population)

length.group.summary.2 <- length.group.summary %>% 
  ungroup() %>%
  relocate(population,.before = "num_group") %>%
  select(population,num_group,num_year_all_group,num_year_group_above_critical) %>%
  gather(key="num_year_criteria",
         value = "num_year",c(num_year_all_group,num_year_group_above_critical))


ggplot(data=length.group.summary.2, aes(x=num_group,y=num_year,color=num_year_criteria))+
  geom_point(size=2)+
  geom_line(aes(group=num_year_criteria))+
  facet_wrap(~population) +
  theme(legend.position="top")
ggsave(paste0("figure/numyear_lengthgroup/cutoff_10/num_year_lengthgroup.tiff"),width = 14,height = 10,dpi=300)



for ( regional.pop in regional.pop.list) {
  
  ggplot(data=length.group.summary[which(length.group.summary$population==regional.pop),],     
       aes(x=yearqua,y=range_group_year,color=length_group))+
  geom_line(aes(group=length_group))+
  geom_point()+
  #geom_smooth(method = "lm",aes(group=Length))+
  facet_wrap(~num_group)
  ggsave(paste0("figure/range_size/cutoff_1/",regional.pop,".tiff"),width = 12,height = 8,dpi=300)
  
  ggplot(data=length.group.summary[which(length.group.summary$population==regional.pop),],     
       aes(x=yearqua,y=log(cpue_group_year),color=length_group))+
  geom_line(aes(group=length_group))+
  geom_point()+
  #geom_smooth(method = "lm",aes(group=Length))+
  facet_wrap(~num_group)
  ggsave(paste0("figure/cpue_lengthgroup/cutoff_1/",regional.pop,".tiff"),width = 12,height = 8,dpi=300)
  
  
}
```



