---
title: "04_Nlengthgroup_Cleaning"
output: html_document
---

This script filters populations out based on criteria 4 (see Methods section of the paper).

#Set working directory
```{r setup}

knitr::opts_knit$set(root.dir = ("/Users/hsiaohang.tao/Dropbox/academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/"))
```

#library
```{r}
library("tidyverse")
library("readxl")
```

#------------Functions--------------------
#<Function>: Reduce the noises

Removing Year with number of lengthgroup < X 
(X depends on Num of lengthgroup)

*Require importing critical.group.num vector (depends on length-group Scenario, can be freely defined)*
Step1. Count the number of length-groups in each year (year.denoise.group)
Step2. Count numyear that acheive critical_lengthgroup or all_lengthgroup
Step3. Merge & Remove Year with number of lengthgroup < X 
```{r}

denoise_group_data <- function(fish.data, group.num, critical.group.num){
  #Step1.
  year.denoise.group <- fish.data %>% 
  ungroup() %>% select(-subarea,-cpue_length_subarea_yearqua) %>%

  group_by(region,species,yearqua) %>%
  summarise(num_group_year=n_distinct(length_group)) %>%
    #Step2.
  group_by(region,species) %>%
  mutate(num_year_group_above_critical=length(which(num_group_year >= critical.group.num)),
         num_year_all_group=length(which(num_group_year==group.num))) %>%
  unique()
  #Step3.
  cpue.length.subarea.yearqua <- left_join(fish.data,year.denoise.group)
  
  cpue.length.subarea.yearqua <- cpue.length.subarea.yearqua %>%
     filter(num_group_year >= critical.group.num)
  
  
  print(paste0("For lengthgroup = ",group.num," senario, subset rows from ", dim(fish.data)[1], " to ",
               dim(cpue.length.subarea.yearqua)[1]))
  
  return(cpue.length.subarea.yearqua)
}

```

#------------Remove noises--------------------
#<With Loop> Remove noises of incomplete lengthgroup
*critical.group.num can be freely defined as we wish*

```{r}
group.num = seq(5,10,1)
cutoff = c(0.1,1,3)

for (i in 1:length(group.num)) {
  for (j in 1:length(cutoff)) {
    
  
    import.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                  "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_cut.csv") 
    all.region.data <- read.csv(import.dir)
    
    import.dir2 = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                  "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/filtered_species_",group.num[i],"_lengthgroup.csv")
    selected.sp <- read.csv(import.dir2) %>% select(region,species)

    all.region.data <- all.region.data %>% inner_join(selected.sp)
  
    all.region.data <- denoise_group_data(all.region.data,
                                          group.num = group.num[i],
                                          critical.group.num = round(0.8*group.num[i]))
  
  
  
    export.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                 "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_denoise.csv")
  
    write.csv(all.region.data, file = export.dir )
  }
}

```

#<Without Loop> Remove noises of incomplete number of lengthgroup
*Run above function step by step*
Step1. Load (1) cpue data with 5 lengthgroup (2) species list
Step2. Filter species with list
Step3. Remove noises & Save data

(demonstrated here with 5 length-groups & 1% cutoff)
```{r}

#Step1.
fish.data <- read.csv("data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_cut.csv")
selected.sp <-  read.csv("data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/filtered_species_5_lengthgroup.csv") %>% select(-X) %>%
  select(region,species)

fish.data <- fish.data %>%
  select(-X) %>% 
  mutate_at("species",as.factor)

#Step2. 
fish.data <- fish.data %>% inner_join(selected.sp)


#Step3.
group.num=5
critical.group.num = 4

cpue.length.subarea.yearqua <- denoise_group_data(fish.data, group.num = 5,
                                                  critical.group.num = 0.8*5)

write.csv(cpue.length.subarea.yearqua,file = "data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_denoise.csv")

```

