---
title: "05_Nlengthgroup_RangeCompiling"
output: html_document
---
#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))

knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))
```

#library
```{r}

library("tidyverse")


```

#------------Functions--------------------
#<Function>: Deriving cpue- and range-related indices for each length
*Length-specific Prop_CPUE/Area_of_Distribution/Prop_Distribution within given species*

Required Input dataset: cpue.length.subarea.yearqua
Here we will create:
(1) cpue_length_yearqua// (2) cpue_year// (3) Prop_CPUE_Length_Year// *Step 1. ~ Step 3.*  
(4) num_subarea_all_length// (5) num_subarea_length// (6) prop_subarea_length *Step 4. ~ Step 7.*


Step1. Create cpue_length_yearqua and Subset cpue.length.subarea.yearqua into Length-Specific data
Step2. Derive cpue_year (Summed CPUE for each species, in each year)
Step3. Derive Prop_CPUE_Length_Year (length-specific relative abundance in each year)

Step4. Derive num_subarea_all_length (Species-specific home range in each year), count subareas with n_distinct()
Step5. Derive num_subarea_length (Length-specific home range in each year) with n_distinct()
Step6. Incorporate num_subarea_all_length & num_subarea_length into cpue.length.year
Step7. Derive prop_subarea_length (Proportional occupation for lengthgroup, given the yearly home range of population)

```{r}
get_length_yearly_data <- function(cpue.length.subarea.yearqua.data){
  
  cpue.length.year <- cpue.length.subarea.yearqua.data %>%
    #Step1.
    group_by(region,species,yearqua,length_group) %>%
    mutate(cpue_length_yearqua=sum(cpue_length_subarea_yearqua)) %>%
    select(region,species, yearqua, length_group, cpue_length_yearqua) %>%
    unique()

  cpue.length.year <- cpue.length.year %>%
    group_by(region,species,yearqua) %>%
    #Step2.
    mutate(cpue_year = sum(cpue_length_yearqua)) %>%
    #Step3.
    mutate(prop_cpue_length_year = cpue_length_yearqua / cpue_year)

    
  #Step4.
  home.range.year <- cpue.length.subarea.yearqua.data %>%
    group_by(region,species, yearqua) %>%
    summarise(num_subarea_all_length=n_distinct(subarea))
  
  #Step5. 
  num.subarea.length <- cpue.length.subarea.yearqua.data %>%
    group_by(region,species, yearqua, length_group) %>%
    summarise(num_subarea_length = n_distinct(subarea)) 

  #Step6.
  cpue.length.year <- left_join(cpue.length.year, num.subarea.length, 
    by = c("region","species","yearqua","length_group"))
  
  cpue.length.year <- left_join(cpue.length.year, home.range.year, 
    by = c("region","species","yearqua"))
 
  #Step7.
  cpue.length.year <- cpue.length.year %>%  
    mutate(prop_subarea_length = num_subarea_length / num_subarea_all_length) %>% 
    mutate(length_group = factor(length_group, levels=c("1","2","3","4","5","6","7","8","9","10")))

  return(cpue.length.year)
  
}

```


#<Function>: Deriving cpue- and range-related indices for species
Required Input dataset: cpue.length.year
*Range-related indices also include Mean Proportional Occupation*

Step1. Create (1) mean_prop_subarea ( arithmetic mean among all length ) 
              (2) mean_num_subarea
Step2. Subset data

```{r}
get_species_yearly_data <- function(cpue.length.year.data){
 
  cpue.year <- cpue.length.year.data %>% 
    group_by(region,species,yearqua) %>%
     #Step1
    mutate(mean_prop_subarea = mean(prop_subarea_length),
           mean_num_subarea = mean(num_subarea_length)) %>%
     #Step2
    select(region,species,yearqua,cpue_year,num_subarea_all_length,
           mean_prop_subarea,mean_num_subarea) %>%
    unique()
  
}

```

#------------Range compilation--------------------
#<With Loop>: Compiling 3 CPUE datasets for all length-groups (5~10)

cpue.length.subarea.yearqua (*Note*)//cpue.length.year//cpue.year

*Run above function with loop*
Step1. Load and arrange cpue.length.subarea.yearqua data
Step2. Compile cpue.length.year
Step3. Compile cpue.year
(suppressed)Step4. Arrange yearqua and Add Temp data to 3 CPUE datasets
Step5. Save data

*Note* The Input data is already sufficient for compiling geographic overlap (see Script 06) and 
the Abundance-weighted range overlap will be compiled in Script 06 
(See Add Intersect Infos & Get Overlap Infos) 
```{r}

group.num <- seq(5,10,1)
cutoff <- c(0.1,1,3)


for (i in 1:length(group.num)) {
    
    for (j in 1:length(cutoff)) {
      
      #Step1.
      import.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                      "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_denoise.csv")
    
      length.group.denoise <- read.csv(import.dir) 
    
      cpue.length.subarea.yearqua <- length.group.denoise %>% 
        select(-X) %>% 
        mutate_at("species",as.factor)

      #Step2.~Step4.

      cpue.length.year <- get_length_yearly_data(cpue.length.subarea.yearqua)
      cpue.year <- get_species_yearly_data(cpue.length.year)

      #Step5.
      export.dir = paste0("data/initial_data/lengthgroup_specific/",group.num[i],
                  "_lengthgroup/with_extreme/cutoff_",cutoff[j],"/",group.num[i],"_lengthgroup_cpue_compiled.RData")

      save(cpue.length.subarea.yearqua, cpue.length.year, cpue.year, file = export.dir)

      print(paste0("cpue data for length-group ",group.num[i]," was finished"))
    }
}
```


*Run above function step by step*
#<without loop>: Compile 3 cpue datasets (5 lengthgroup case)
Step1. Load and arrange cpue.length.subarea.yearqua data
Step2. Compile cpue.length.year
Step3. Compile cpue.year
```{r}
#Step1.
length.group.denoise <- read.csv("data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_denoise.csv")

cpue.length.subarea.yearqua <- length.group.denoise %>% 
  select(-X) %>% 
  mutate_at("species",as.factor)

#Step2.
cpue.length.year <- get_length_yearly_data(cpue.length.subarea.yearqua)

#Step3.
cpue.year <- get_species_yearly_data(cpue.length.year)
```















