---
title: "06_Nlengthgroup_OverlapCalculating"
output: html_document
---
#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))

knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))
```

library
```{r}

library("tidyverse")


```

#------------Functions--------------------
#<Function>: Get subarea infos for Length Pair
*Required Input Dataset: cpue.length.subarea.yearqua*

step1. calculate cpue_length_yearqua
step2. select columns & filter for target length group for given pair 
step3. count lengthgroup-specific range_size in each year
step4. rename columns with specific ID within given pair

*var: names waiting for being changed to pair-specific ID format
```{r}
get_lengthgroup_subarea_infos <- function(cpue.length.subarea.yearqua, group.num, pair.id){
    
  var <- c("pair_id","cpue_length_subarea_yearqua","cpue_length_yearqua","range_size")
  pair.id.append <- ifelse(pair.id=="x","_x","_y")
  
  cpue.group <- cpue.length.subarea.yearqua %>%
      #step1.
      group_by(region,species,yearqua,length_group) %>%
      mutate(cpue_length_yearqua=sum(cpue_length_subarea_yearqua)) %>%
      relocate(cpue_length_yearqua,.after="cpue_length_subarea_yearqua") %>%
      #step2.
      select(region,species, yearqua, subarea, 
             length_group, cpue_length_subarea_yearqua, cpue_length_yearqua) %>%
      group_by(region,species,yearqua) %>%                 
      filter(length_group == group.num & cpue_length_subarea_yearqua>0) %>%
      #step3.
      mutate(range_size = n_distinct(subarea)) %>%
      #step4.
      rename("pair_id" = "length_group") %>% 
      rename_with(.fn = ~paste0(., pair.id.append), .cols = all_of(var))
   
  return(cpue.group)
}


```


#<Function>: get union infos
required input dataset: cpue.group.x & cpue.group.y

step1. derive both.groups.exist.year, then discard year records where only 1 group was surveyed
       (acheived by taking inner_join between cpue.group and both.groups.exist.year)
       
step2. take distribution unions, then count union subareas during both.groups.exist.year  
step3. return data (depends on pair.id)

```{r}
get_union_infos <- function(cpue.group.x,cpue.group.y,pair.id){
  #step1.
  subset.var <-c("region","species","yearqua")
  both.groups.exist.year <- intersect(cpue.group.x[,subset.var],cpue.group.y[,subset.var])

  #@HHT add by = c("region", "species", "yearqua")
  cpue.group.x <- inner_join(both.groups.exist.year,cpue.group.x,by = c("region", "species", "yearqua"))
  cpue.group.y <- inner_join(both.groups.exist.year,cpue.group.y,by = c("region", "species", "yearqua"))

  #step2.
  subset.var2 <-c("region","species","yearqua","subarea")
  union <- union(cpue.group.x[,subset.var2], cpue.group.y[,subset.var2]) %>%        
    #@HHT 2023.05.30 I added .group='drop" to mute error message
    group_by (region,species,yearqua) %>%                                  #
    summarise(num_union=n_distinct(subarea),.groups = 'drop')
  
  #step3.
  
  #@HHT I added by = c("region", "species", "yearqua") to mute error message
  if(pair.id == "x"){
     cpue.group.x <- left_join(cpue.group.x, union,
                               by = c("region", "species", "yearqua"))
    return(cpue.group.x)
  }
  
  if(pair.id == "y"){
    cpue.group.y <- left_join(cpue.group.y, union,by = c("region", "species", "yearqua"))
    return(cpue.group.y)
  }
  
}

```


#<Function>: get intersect infos
step1. get subarea records within cpue.group.x & cpue.group.y during both.groups.exist.year via full_join
*note that in some years, both groups shared no overlapping subarea (intersect = 0), although they did exist*
*some values would become NA after taking full_join, since two groups did not co-occur there.*
*so, in "step.3" we need to replace the NA values with true values showing in cpue.group.x/y*

step2. check whether both groups co-occurred in same subareas (overlap_region = 1 or 0) 

step3. replace NA into correct values in each column
       step3-(1) pair_id_x/y
       step3-(2) cpue_length_subarea_yearqua_x/y
       step3-(3) cpue_length_yearqua_x/y & range_size_x/y
      
step4. derive num_intersect & cpue_length_intersect_yearqua_x/y
step5. reorder columns
```{r}
get_intersect_infos <- function(cpue.group.x,cpue.group.y,i){
  #step1
  cpue.group.xy <- full_join(cpue.group.x,cpue.group.y,
                       by=c("region","species","yearqua","subarea","num_union"))
  #step2
  cpue.group.xy$overlap_region <- ifelse(is.na(cpue.group.xy$cpue_length_subarea_yearqua_x)|
                                         is.na(cpue.group.xy$cpue_length_subarea_yearqua_y), 0, 1)
  
  #step3-(1)
  cpue.group.xy$pair_id_x <- combn(1:num.combn, 2)[,i][1]
  cpue.group.xy$pair_id_y <- combn(1:num.combn, 2)[,i][2]
  
  #step3-(2)
  cols <- c("cpue_length_subarea_yearqua_x","cpue_length_subarea_yearqua_y")
  cpue.group.xy[cols][is.na(cpue.group.xy[cols])] <- 0
  #step3-(3)
  cpue.group.xy <- cpue.group.xy %>%
    group_by(region,species,yearqua) %>%
    mutate_at(vars(cpue_length_yearqua_x,cpue_length_yearqua_y,range_size_x,range_size_y),
              ~ifelse(all(is.na(.))==TRUE,0,na.omit(.)[1]))
  
  ##step4
  cpue.group.xy <- cpue.group.xy %>%
    group_by(region,species,yearqua) %>%
    mutate(num_intersect=sum(overlap_region),
           cpue_length_intersect_yearqua_x= 
            ifelse(num_intersect==0,0,sum(cpue_length_subarea_yearqua_x[which(overlap_region==1)])),
           cpue_length_intersect_yearqua_y= 
            ifelse(num_intersect==0,0,sum(cpue_length_subarea_yearqua_y[which(overlap_region==1)])))
  #step5
  cpue.group.xy <- cpue.group.xy %>%
    relocate(overlap_region,.after = "subarea") %>%
    relocate(num_intersect,.before = "num_union") %>%
    relocate(cpue_length_intersect_yearqua_x,.before = "cpue_length_yearqua_x") %>%
    relocate(cpue_length_intersect_yearqua_y,.before = "cpue_length_yearqua_y")
  
  return(cpue.group.xy)
  
}
```


#<function>: get overlap indices
step1. derive precursors for overlap indices:
       prop_cpue_intersect_x/y & 
       prop_cpue_length_subarea_yearqua_x/y/xy &
       sum_prop_cpue_length_subarea_yearqua_x/y_square
       
step2. derive overlap indices:
       area_overlap/overlap_abun_weighted (regional scale indices)
       overlap_bhat/schoener/local_collocation (grid-based indices)
       
step3. remove unecessary variables and subset dataframe

```{r}

get_overlap_indices <- function(cpue.group.xy){
  #step1
  cpue.group.xy <- cpue.group.xy %>%
    mutate(prop_cpue_intersect_x = cpue_length_intersect_yearqua_x / cpue_length_yearqua_x,
           prop_cpue_intersect_y = cpue_length_intersect_yearqua_y / cpue_length_yearqua_y) #%>%
    #mutate(prop_cpue_length_subarea_yearqua_x = cpue_length_subarea_yearqua_x / cpue_length_yearqua_x,
    #      prop_cpue_length_subarea_yearqua_y = cpue_length_subarea_yearqua_y / cpue_length_yearqua_y,
    #      prop_cpue_length_subarea_yearqua_xy = sqrt(prop_cpue_length_subarea_yearqua_x *        
    #                                                 prop_cpue_length_subarea_yearqua_y)) %>%
    #group_by(region,species,yearqua) %>%
    #mutate(sum_prop_cpue_length_subarea_yearqua_x_square = sum(prop_cpue_length_subarea_yearqua_x^2),
    #       sum_prop_cpue_length_subarea_yearqua_y_square = sum(prop_cpue_length_subarea_yearqua_y^2))
    
    #step2
    cpue.group.xy <- cpue.group.xy %>%
    mutate(area_overlap =  num_intersect / num_union,
          overlap_abun_weighted = sqrt(prop_cpue_intersect_x * prop_cpue_intersect_y)) #%>%
      
    #group_by(region,species,yearqua) %>%
    #mutate(overlap_bhat = sum(prop_cpue_length_subarea_yearqua_xy),
    #      overlap_schoener=1-sum(0.5*abs(prop_cpue_length_subarea_yearqua_x-prop_cpue_length_subarea_yearqua_y)),
     #     overlap_local_collocation=sum(prop_cpue_length_subarea_yearqua_xy^2) / 
     #      sqrt(sum_prop_cpue_length_subarea_yearqua_x_square*sum_prop_cpue_length_subarea_yearqua_y_square)) 

  #step3
  cpue.group.xy <- cpue.group.xy %>%
    ungroup() %>%
    select(-subarea,-overlap_region,
           -cpue_length_subarea_yearqua_x,-cpue_length_subarea_yearqua_y#,
           #-prop_cpue_length_subarea_yearqua_x,-prop_cpue_length_subarea_yearqua_y,
           #-prop_cpue_length_subarea_yearqua_xy,
           #-sum_prop_cpue_length_subarea_yearqua_x_square,
           #-sum_prop_cpue_length_subarea_yearqua_y_square
           ) %>%
    unique()
  
  return(cpue.group.xy)
}
```

#<Function>: get species-level mean overlap infos by averaging all paired ovelap infos
```{r}

get_species_level_mean_overlap<- function(overlap.infos.all.pair){
  
  overlap.infos.all.pair <- overlap.infos.all.pair %>% 
    group_by(region,species,yearqua) %>%
    mutate(mean_intersect = mean(num_intersect),
           mean_union = mean(num_union),
           mean_oa = mean(area_overlap),
           mean_ow = mean(overlap_abun_weighted)#,
           #mean_overlap_bhat = mean(overlap_bhat),
           #mean_overlap_schoener=mean(overlap_schoener),
           #mean_overlap_local_collocation=mean(overlap_local_collocation)
           ) 
    
    return(overlap.infos.all.pair)
}

```

#<Function>: add_overlap_infos_to_cpue_year (Add Species-level mean overlap index to cpue.year)
```{r}

add_overlap_infos_to_cpue_year <- function(overlap.infos.all.pair,cpue.year){

    overlap.infos.all.pair <- overlap.infos.all.pair %>%
      select(region,species,yearqua,
             mean_intersect,mean_union,
             mean_oa,mean_ow) %>%
      group_by(region,species,yearqua) %>%
      unique()

    cpue.year <- inner_join(overlap.infos.all.pair, cpue.year, 
                    by = c("region","species","yearqua"))
    
    return(cpue.year)

}

```

#<Function>: save cpue.year/cpue.pair
```{r}

save_cpue_year_cpue_pair <- function(cpue.year,cpue.pair,group.num,cutoff){
  
  save(cpue.year,cpue.pair,
       file = paste0("data/overlap_data/lengthgroup_specific/",group.num,
                     "_lengthgroup/with_extreme/cutoff_",cutoff,"/",group.num,"_lengthgroup_yearly_overlap_infos.RData"))
  
  print(paste0(group.num,"_length_group_yearly_overlap_infos.RData","  is saved."))
  
}
```

#------------Overlap compilation--------------------
*Run above function with loop*
#<With Loop>: Compile overlap data (cpue.year/cpue.pair)

@HHT 2022.03.23----------------------
I have many messages: 
`summarise()` has grouped output by 'region', 'species'. You can override using the `.groups` argument.
Joining, by = c("region", "species", "yearqua")

This prevents from running the script. Maybe I need to update my R or package---------


-----
`summarise()` has grouped output by 'region', 'species'. You can override using the `.groups` argument.
Joining, by = c("region", "species", "yearqua")
Joining, by = c("region", "species", "yearqua")
Joining, by = c("region", "species", "yearqua")
`summarise()` has grouped output by 'region', 'species'. You can override using the `.groups` argument.
Joining, by = c("region", "species", "yearqua")
-----


```{r}
group.num <- seq(5,10,1)
cutoff <- c(0.1,1,3)

for ( code in 1:length(group.num)) {
  
  for ( k in 1:length(cutoff)) {
  
  load(paste0("data/initial_data/lengthgroup_specific/",group.num[code],"_lengthgroup/with_extreme/cutoff_",
              cutoff[k],"/",group.num[code],"_lengthgroup_cpue_compiled.RData"))
  
  num.combn = group.num[code]  

  cpue.pair.output <- vector(mode = "list")
  
  for (i in 1 : ncol(combn(1:num.combn,2)) ){
    
    x <- combn(1:num.combn, 2)[,i][1]
    y <- combn(1:num.combn, 2)[,i][2]
    
    cpue.group.x <- get_lengthgroup_subarea_infos(cpue.length.subarea.yearqua, 
                                                   group.num = x, pair.id = "x")
    cpue.group.y <- get_lengthgroup_subarea_infos(cpue.length.subarea.yearqua, 
                                                   group.num = y, pair.id = "y")
    
    cpue.group.x <- get_union_infos(cpue.group.x,cpue.group.y, pair.id = "x")
    cpue.group.y <- get_union_infos(cpue.group.x,cpue.group.y, pair.id = "y")
    
    cpue.group.xy <- get_intersect_infos(cpue.group.x,cpue.group.y,i)
    
    cpue.group.xy <- get_overlap_indices(cpue.group.xy)

    cpue.pair.output[[i]] <- cpue.group.xy
  
    
}  

    overlap.infos.all.pair <- bind_rows(cpue.pair.output,.id = "pair_id")
    
    overlap.infos.all.pair <- get_species_level_mean_overlap(overlap.infos.all.pair)
    
    cpue.year <- add_overlap_infos_to_cpue_year(overlap.infos.all.pair, cpue.year)
    cpue.pair <- overlap.infos.all.pair
    save_cpue_year_cpue_pair(cpue.year, cpue.pair, group.num[code],cutoff[k])
    
  }

}
```

*Run above function step by step*
#<without loop>: Compile overlap data (cpue.year/cpue.pair) (5 lengthgroup case)
```{r}
group.num = 5
cutoff = 1

load(paste0("data/initial_data/lengthgroup_specific/",group.num,"_lengthgroup/with_extreme/cutoff_",cutoff,"/",group.num,"_lengthgroup_cpue_compiled.RData"))

num.combn = group.num

cpue.pair.output <- vector(mode = "list")

for (i in 1 : ncol(combn(1:num.combn,2)) ){
    
    x <- combn(1:num.combn, 2)[,i][1]
    y <- combn(1:num.combn, 2)[,i][2]
    
    cpue.group.x <- get_lengthgroup_subarea_infos(cpue.length.subarea.yearqua, 
                                                   group.num = x, pair.id = "x")
    cpue.group.y <- get_lengthgroup_subarea_infos(cpue.length.subarea.yearqua, 
                                                   group.num = y, pair.id = "y")
    
    cpue.group.x <- get_union_infos(cpue.group.x,cpue.group.y, pair.id = "x")
    cpue.group.y <- get_union_infos(cpue.group.x,cpue.group.y, pair.id = "y")
    
    cpue.group.xy <- get_intersect_infos(cpue.group.x,cpue.group.y,i)
    
    cpue.group.xy <- get_overlap_indices(cpue.group.xy)

    cpue.pair.output[[i]] <- cpue.group.xy

}  

    overlap.infos.all.pair <- bind_rows(cpue.pair.output,.id = "pair_id") %>%
      arrange(region,species,pair_id,yearqua)
    overlap.infos.all.pair <- get_species_level_mean_overlap(overlap.infos.all.pair)
    cpue.year <- add_overlap_infos_to_cpue_year(overlap.infos.all.pair, cpue.year) %>%
      arrange(region,species,yearqua)
    cpue.pair <- overlap.infos.all.pair
    #save_cpue_year_cpue_pair(cpue.year, cpue.pair, group.num,cutoff)
    
    cpue.pair %>% filter(num_intersect==0) %>%
    ungroup() %>%
    group_by(species) %>%
    summarise(num_pair_no_intersect=n())
    
```










