---
title: "coocurrence"
output: html_document
date: "2025-01-10"
editor_options: 
  chunk_output_type: inline
---

This script calculates co-occurrence index Alpha-MLE.

#Set working directory
```{r setup}

knitr::opts_knit$set(root.dir = ("/Users/hsiaohang.tao/Dropbox/academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/"))

```



# library
```{r}
library(dplyr)
library(tidyverse)
library(CooccurrenceAffinity)
library(cooccur)
library(data.table)
library(mapplots)
```




# read data
```{r}
getwd()

dd <- read.csv("output/cpue.length.subarea.yearqua.csv")


# check data length of Baltic Sea
cpue.length.subarea.yearqua %>% filter(region == "Baltic Sea", species == "Clupea harengus") %>% 
  group_by(yearqua) %>% 
  tally()

```



# create function to calculate Alpha MLE 
In the function, each region is cut to two halves and run separately to accommodate comutation power 

split to each region and convert to wide form so that N for each region is specific

consider to assign N = max(ma,mb)
Or split to each region + species
so N for each region and species is specific.
Otherwise, for example in north sea, N is always 198
while the spatial range of a species can be much less limited

```{r}

co_fun <- function(data, region_name, divide){
  d1 <- dd %>% 
  dplyr::select(region, species, yearqua, subarea,length_group, cpue_length_subarea_yearqua) 

d1 <- d1 %>% 
  filter(region == region_name) %>% 
  group_by(species, yearqua, length_group) %>% 
  pivot_wider(names_from = subarea,
              values_from = 
                cpue_length_subarea_yearqua) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) %>%
  replace(is.na(.), 0) %>% 
  ungroup() %>% 
  group_by(region, species, yearqua) %>% 
  mutate_at('length_group', as.factor)
  
d.list <- group_split(d1)

# convert to dataframe
d.list <- map(d.list, as.data.frame)

# test by 4 lists
#d.list <- d.list[c(1:4)]

# subset list by two
if (divide == 1){d.list.sub <- d.list[c(1:(length(d.list)/2))]}else{
  d.list.sub <- d.list[c(((length(d.list)/2)+1):length(d.list))]
}

# loop
out <- list(NA)

for (i in 1:length(d.list.sub)){
  out[i] <- affinity(data = d.list.sub[[i]][,5:ncol((d.list.sub)[[i]])], 
                       row.or.col = "row")[1]
  }
print(out)
}

```


# run co-occurence function
```{r}

# run co_fun 
out.ns.1 <- co_fun(dd, 'North Sea', 1)
out.ns.2 <- co_fun(dd, 'North Sea', 2)
out.swc.1 <- co_fun(dd, 'Scottish West Coast', 1)
out.swc.2 <- co_fun(dd, 'Scottish West Coast', 2)
out.ebs.1 <- co_fun(dd, 'Eastern Bering Sea', 1)
out.ebs.2 <- co_fun(dd, 'Eastern Bering Sea', 2)
out.bs.1 <- co_fun(dd, 'Baltic Sea', 1)
out.bs.2 <- co_fun(dd, 'Baltic Sea', 2)

# check
out.ns.1[[1]]

# bind two halves for each region
out.ns <- c(out.ns.1, out.ns.2)
out.swc <- c(out.swc.1, out.swc.2)
out.ebs <- c(out.ebs.1, out.ebs.2)
out.bs <- c(out.bs.1, out.bs.2)


```


# house-keeping
```{r}

# create tibble of region, species, yearqua
d.name <- function(region_name) {dd %>% 
  separate(yearqua, into = c("year", "qua")) %>% 
  dplyr::select(region, species, year) %>%
  filter(region == region_name) %>% 
  unique()}

ns.name <- as.data.frame(d.name('North Sea'))
swc.name <- as.data.frame(d.name('Scottish West Coast'))
ebs.name <- as.data.frame(d.name('Eastern Bering Sea'))
bs.name <- as.data.frame(d.name('Baltic Sea'))

# bind name tibble with co-occurrence outputs
for(i in 1:length(out.ns)){
out.ns[[i]]$region <- ns.name[i,1]
out.ns[[i]]$species <- ns.name[i,2]
out.ns[[i]]$year <- ns.name[i,3]}

for(i in 1:length(out.swc)){
out.swc[[i]]$region <- swc.name[i,1]
out.swc[[i]]$species <- swc.name[i,2]
out.swc[[i]]$year <- swc.name[i,3]}

for(i in 1:length(out.ebs)){
out.ebs[[i]]$region <- ebs.name[i,1]
out.ebs[[i]]$species <- ebs.name[i,2]
out.ebs[[i]]$year <- ebs.name[i,3]}

for(i in 1:length(out.bs)){
out.bs[[i]]$region <- bs.name[i,1]
out.bs[[i]]$species <- bs.name[i,2]
out.bs[[i]]$year <- bs.name[i,3]}



# in each list,bind all dataframes to one dataframe
out.ns <- data.table::rbindlist(out.ns) %>%
  dplyr::as_data_frame()

out.swc <- data.table::rbindlist(out.swc) %>%
  dplyr::as_data_frame()

out.ebs <- data.table::rbindlist(out.ebs) %>%
  dplyr::as_data_frame()

out.bs <- data.table::rbindlist(out.bs) %>%
  dplyr::as_data_frame()


# bind four dataframes
dd.co <- data.table::rbindlist(list(out.ns, out.swc,out.ebs,out.bs)) %>%
  dplyr::as_data_frame()

dd.co <- as_tibble(dd.co)

# rename
dd.co <- dd.co %>% rename(size_group_1 = entity_1,
                 size_group_2 = entity_2) %>% 
  relocate(region, .before = size_group_1) %>% 
   relocate(species,.after = region) %>% 
   relocate(year, .after = species) %>% 
  mutate(size_group_1 = str_remove(size_group_1, "V"),
         size_group_2 = str_remove(size_group_2, "V"))
 
```


# save & read dd.co
```{r}         

write.csv(dd.co,"output/co.dd.csv",row.names=F)
dd.co <- read.csv("output/co.dd.csv")

```


