---
title: "coocurrence"
output: html_document
date: "2025-01-10"
editor_options: 
  chunk_output_type: inline
---

# log

20240113
calculate alpha by running the function by region, but may consider to also run by species. Next: now the memory for running is too much, I may try to subset the outputs.
20250117 
til line 223 I calculated alpha for all species for each region, pairs, years. Next is to combine all lists, then analyze, perhaps follow 07_plot_v4.
20250118
Completed to save alpha as <dd.co>

# library
```{r}
library(dplyr)
library(tidyverse)
library(CooccurrenceAffinity)
library(cooccur)
library(data.table)
library(mapplots)
```


#Set working directory
```{r setup}

knitr::opts_knit$set(root.dir = ("/Users/hsiaohang.tao/Dropbox/Academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/"))

getwd()
```


# read data
```{r}
getwd()

load("/Users/hsiaohang.tao/Dropbox/Academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/data/initial_data/lengthgroup_specific/5_lengthgroup/with_extreme/cutoff_1/5_lengthgroup_cpue_compiled.RData")

dd <- cpue.length.subarea.yearqua

# check data length of Baltic Sea
cpue.length.subarea.yearqua %>% filter(region == "Baltic Sea",
                                       species == "Clupea harengus") %>% 
  group_by(yearqua) %>% 
  tally()

```


# try function for one population
```{r}

dd <- cpue.length.subarea.yearqua

#baltic_sea
d.trial <- dd %>% 
  filter(region == 'Eastern Bering Sea') %>% 
  filter(yearqua == 2000) %>% 
  dplyr::select(species, yearqua, subarea,length_group, cpue_length_subarea_yearqua) %>% 
  filter(species == 'Atheresthes stomias') %>% 
  select(subarea, length_group, cpue_length_subarea_yearqua) %>% 
    mutate_at('length_group', as.factor) %>% 
  pivot_wider(names_from = subarea,
              values_from = 
                cpue_length_subarea_yearqua) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) %>%
  replace(is.na(.), 0) %>% 
  remove_rownames %>% column_to_rownames(var="length_group")

d.trial

affinity(data = d.trial, row.or.col = "row")

```


# run all data
This has problem as all subareas from all regions are shown in the columns after converting to wider format
```{r, eval =FALSE}

# split data to list <d.list>
d1 <- dd %>% 
  dplyr::select(region, species, yearqua, subarea,length_group, cpue_length_subarea_yearqua) %>% 
  group_by(region, species, yearqua, length_group) %>% 
  pivot_wider(names_from = subarea,
              values_from = 
                cpue_length_subarea_yearqua) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) %>%
  replace(is.na(.), 0) %>% 
  ungroup() %>% 
  group_by(region, species, yearqua) %>% 
  mutate_at('length_group', as.factor)

d.list <- group_split(d1)
d.list[[1]]


# convert to dataframe
d.list.dataframe <- map(d.list, as.data.frame)

d.list.dataframe[[2]]

d.list.demo <- list(d.list.dataframe[[100]],d.list.dataframe[[101]])


# loop
out <- list(NA)

for (i in 1:length(d.list.demo)){
  out[[i]] <- affinity(data = d.list.demo[[i]][,5:ncol((d.list.demo)[[i]])], 
                       row.or.col = "row")
}

out[[1]][1]

```


# run all data by region: snippet
split to each region and convert to wide form
so N for each region is specific

consider to assign N = max(ma,mb)
Or split to each region + species
so N for each region and species is specific.
Otherwise, for example in north sea, N is always 198
while the spatial range of a species can be much less limited

```{r}

d1 <- dd %>% 
  dplyr::select(region, species, yearqua, subarea,length_group, cpue_length_subarea_yearqua) 

d1 <- d1 %>% 
  filter(region == 'North Sea') %>% 
  group_by(species, yearqua, length_group) %>% 
  pivot_wider(names_from = subarea,
              values_from = 
                cpue_length_subarea_yearqua) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) %>%
  replace(is.na(.), 0) %>% 
  ungroup() %>% 
  group_by(region, species, yearqua) %>% 
  mutate_at('length_group', as.factor)
  
d.list <- group_split(d1)
d.list[[1]]

# convert to dataframe
d.list <- map(d.list, as.data.frame)


# cut into two lists 1026 / 2 = 513
d.list.1 <- d.list[c(1:(length(d.list)/2))]
length(d.list)/2

# loop
out.ns.1 <- list(NA)

for (i in 1:length(d.list.1)){
  out.ns.1[[i]] <- affinity(data = d.list[[i]][,5:ncol((d.list)[[i]])], 
                       row.or.col = "row")
}

out.ns.1[[2]]


```


# co_function
In the function, each region is cut to two halves and run separately to accommodate comutation power 
```{r}

co_fun <- function(data, region_name, divide){
  d1 <- dd %>% 
  dplyr::select(region, species, yearqua, subarea,length_group, cpue_length_subarea_yearqua) 

d1 <- d1 %>% 
  filter(region == region_name) %>% 
  group_by(species, yearqua, length_group) %>% 
  pivot_wider(names_from = subarea,
              values_from = 
                cpue_length_subarea_yearqua) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) %>%
  replace(is.na(.), 0) %>% 
  ungroup() %>% 
  group_by(region, species, yearqua) %>% 
  mutate_at('length_group', as.factor)
  
d.list <- group_split(d1)

# convert to dataframe
d.list <- map(d.list, as.data.frame)

# test by 4 lists
#d.list <- d.list[c(1:4)]

# subset list by two
if (divide == 1){d.list.sub <- d.list[c(1:(length(d.list)/2))]}else{
  d.list.sub <- d.list[c(((length(d.list)/2)+1):length(d.list))]
}

# loop
out <- list(NA)

for (i in 1:length(d.list.sub)){
  out[i] <- affinity(data = d.list.sub[[i]][,5:ncol((d.list.sub)[[i]])], 
                       row.or.col = "row")[1]
  }
print(out)
}

```

# run co-occurence function
```{r}

# run co_fun 
out.ns.1 <- co_fun(dd, 'North Sea', 1)
out.ns.2 <- co_fun(dd, 'North Sea', 2)
out.swc.1 <- co_fun(dd, 'Scottish West Coast', 1)
out.swc.2 <- co_fun(dd, 'Scottish West Coast', 2)
out.ebs.1 <- co_fun(dd, 'Eastern Bering Sea', 1)
out.ebs.2 <- co_fun(dd, 'Eastern Bering Sea', 2)
out.bs.1 <- co_fun(dd, 'Baltic Sea', 1)
out.bs.2 <- co_fun(dd, 'Baltic Sea', 2)

# check
out.ns.1[[1]]

# bind two halves for each region
out.ns <- c(out.ns.1, out.ns.2)
out.swc <- c(out.swc.1, out.swc.2)
out.ebs <- c(out.ebs.1, out.ebs.2)
out.bs <- c(out.bs.1, out.bs.2)


```

# house-keeping
```{r}

# create tibble of region, species, yearqua
d.name <- function(region_name) {dd %>% 
  separate(yearqua, into = c("year", "qua")) %>% 
  dplyr::select(region, species, year) %>%
  filter(region == region_name) %>% 
  unique()}

ns.name <- as.data.frame(d.name('North Sea'))
swc.name <- as.data.frame(d.name('Scottish West Coast'))
ebs.name <- as.data.frame(d.name('Eastern Bering Sea'))
bs.name <- as.data.frame(d.name('Baltic Sea'))

# bind name tibble with co-occurrence outputs
for(i in 1:length(out.ns)){
out.ns[[i]]$region <- ns.name[i,1]
out.ns[[i]]$species <- ns.name[i,2]
out.ns[[i]]$year <- ns.name[i,3]}

for(i in 1:length(out.swc)){
out.swc[[i]]$region <- swc.name[i,1]
out.swc[[i]]$species <- swc.name[i,2]
out.swc[[i]]$year <- swc.name[i,3]}

for(i in 1:length(out.ebs)){
out.ebs[[i]]$region <- ebs.name[i,1]
out.ebs[[i]]$species <- ebs.name[i,2]
out.ebs[[i]]$year <- ebs.name[i,3]}

for(i in 1:length(out.bs)){
out.bs[[i]]$region <- bs.name[i,1]
out.bs[[i]]$species <- bs.name[i,2]
out.bs[[i]]$year <- bs.name[i,3]}



# in each list,bind all dataframes to one dataframe
out.ns <- data.table::rbindlist(out.ns) %>%
  dplyr::as_data_frame()

out.swc <- data.table::rbindlist(out.swc) %>%
  dplyr::as_data_frame()

out.ebs <- data.table::rbindlist(out.ebs) %>%
  dplyr::as_data_frame()

out.bs <- data.table::rbindlist(out.bs) %>%
  dplyr::as_data_frame()


# bind four dataframes
dd.co <- data.table::rbindlist(list(out.ns, out.swc,out.ebs,out.bs)) %>%
  dplyr::as_data_frame()

dd.co <- as_tibble(dd.co)

# rename
dd.co <- dd.co %>% rename(size_group_1 = entity_1,
                 size_group_2 = entity_2) %>% 
  relocate(region, .before = size_group_1) %>% 
   relocate(species,.after = region) %>% 
   relocate(year, .after = species) %>% 
  mutate(size_group_1 = str_remove(size_group_1, "V"),
         size_group_2 = str_remove(size_group_2, "V"))
 
```

# save & read dd.co
```{r}         

file.name.co <- "/Users/hsiaohang.tao/Dropbox/Academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/data/co.dd.csv"

write.csv(dd.co,file.name.co,row.names=F)
dd.co <- read.csv(file.name.co)

```


# check Baltic Sea data
herring: Clupea harengus (slighter bigger fish)
sprat: Sprattus sprattus (the smallest fish)

Both species in the Baltic Sea has so high overlapping among each pair so alpha is at the ceiling.

In the North Sea, sprat also has high overlapping and high alpha. But it is not yet to the ceiling.
```{r}
dd.co %>% 
  filter(region == 'Baltic Sea')

dd.co %>% 
  filter(region == 'North Sea' &
           species == 'Sprattus sprattus') 

# check proportion of co-occurred N/total N
dd.co %>% 
  filter(region == 'Baltic Sea') %>% 
  mutate(prop_n = obs_cooccur_X/total_N) %>% 
  dplyr::select(prop_n) %>% 
  summarise(median = median(prop_n),
            min = min(prop_n),
            max = max(prop_n),
            mean = mean(prop_n))
  

```



# check Baltic Sea with map: background
lims(alpha = c(-3,10)) to make the overall color darker
Export:result in console-->export-->save as PDF-->
Device 14 x 21 inches, landscape.

Baltic sea geo-coordinates:
https://marineregions.org/gazetteer.php?p=details&id=2401

```{r}

# plot Baltic Sea background

xmin <- 9
xmax <- 30
ymin <- 52
ymax <- 67

background <- ggplot()+
  borders("world", xlim = c(xmin, xmax), ylim = c(ymin, ymax), fill = "grey70",colour = NA, size = 4) +
  scale_size_area() +
  coord_map(xlim = c(xmin, xmax), ylim = c(ymin, ymax))+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), 
    axis.line=element_line(colour = "black"),
    #axis.text.x=element_text(size = 22),
    #axis.text.y = element_text(size = 22),
    #axis.ticks.x=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    #legend.position="none",
    ##legend.title = element_blank will remove all title
    legend.title = element_blank(),
    ##strip.text.x = element_blank() will remove x axis text
    strip.text.x = element_blank(),
    ## Change legend font size
    legend.text=element_text(size=22))+
  ## This change overall legend text and symbol sizes
  guides(alpha = guide_legend(override.aes = list(size = 5)))

background

```

# check Baltic Sea with map: size groups
```{r}


# Filter by species name
d1 <- cpue.length.subarea.yearqua %>% 
  filter(species == 'Clupea harengus' &
         region =='Baltic Sea' &
           length_group %in% c("1","2")) %>% 
    filter (yearqua %in% c("1984","2000")) %>% 
  mutate(lon = ices.rect(subarea)$lon,
         lat = ices.rect(subarea)$lat) %>% 
  mutate_at('yearqua', as.character) %>% 
  mutate_at('length_group', as.character)
  
  
# Plot small + big
background +  
  geom_point(shape = 15, alpha = 0.5,
             aes(x = lon, y = lat, size = 0.5,  
                 color= length_group), data = d1)+
  facet_wrap(~yearqua,ncol=2)+
  scale_color_manual (values = c("red","blue"),
                      name = "length_group ", 
                      labels = c("1", "2"))+
  guides(color = guide_legend(override.aes = list(size = 5)))
  

cpue.length.subarea.yearqua

```



#-------affinity demo----------- 

#function of co-occurence
```{r}

head(finches) 
myout <- affinity(data = finches, row.or.col = "col")

# you can simply flip the analysis to rows to compute species-pair affinity
myout1 <- affinity(data = finches, row.or.col = "row")


```


# variations
```{r}
# the rows of the outputs above include every single pair of the entities,
# producing many columns for various quantities.
# for several of these columns, the function allows outputing the square NxN matrix of the entities
# an example is here
myout2 <- affinity(data = finches, row.or.col = "col", squarematrix = c("alpha_mle", "jaccard"))

# you can also compute all the square matrices with an "all"
myout3 <- affinity(data = finches, row.or.col = "col", squarematrix = c("all"))

# when you want to compute for only certain pairs
myout4 <- affinity(data = finches, row.or.col = "col", which.row.or.col = 4:6,
                  squarematrix = c("alpha_mle"))

myout5 <- affinity(data = finches, row.or.col = "col",
                  which.row.or.col = c("Isabella", "Espanola"), squarematrix = c("alpha_mle"))


# you can also add additional arguments bringing them from ML.Alpha() or AlphInts()
myout5 <- affinity(data = finches, row.or.col = "col",
                   which.row.or.col = c("Isabella", "Espanola"), lev=0.95, pvalType="Blaker")

myout6 <- affinity(data = finches, row.or.col = "col",
                   which.row.or.col = c("Isabella", "Espanola"), lev=0.90, pvalType="Blaker")

# myout5 and myout6 were generated with identical arguments except a difference in "lev",
# which gave different confidence intervals

myout7 <- affinity(data = finches, row.or.col = "col",
                   which.row.or.col = 4:6, lev=0.95, pvalType="Blaker")

myout8 <- affinity(data = finches, row.or.col = "col",
                   which.row.or.col = 4:6, lev=0.95, pvalType="midP")


```


# convert data format
```{r}

# when you have abundance data requiring conversion to binary
# -----------------------------------------------------------
# abundance data is converted to binary based on a threshold supplied.
# it might be a good idea to explore dataprep() function and its examples
# first before workign on affinity() for abundance data.
matrix.data <- matrix(runif(400, 0, 10), nrow = 100, ncol = 4)
row.names(matrix.data) <- paste0("id_", 1:nrow(matrix.data))
colnames(matrix.data) <- paste0("variable_", 1:ncol(matrix.data))

# add some missing data and zero abundance
matrix.data[1,1] <- matrix.data[2,3] <- matrix.data[1,4] <- matrix.data[1,2] <- NA
matrix.data[10,4] <- 0
head(matrix.data)



# now this is an abundance data with some missing and some zero occurrences

# inspecting how the abundance is converted to binary first
dataprep(data = matrix.data, row.or.col = "col", datatype = "abundance",
         threshold = 5, class0.rule = "less")


```

#------------------------------------