---
title: "plot_cooccurence"
output: html_document
editor_options: 
  chunk_output_type: inline
---

This script generates all the graphs in the paper.

#----- Call data & house keeping

# Setup
```{r setup}

knitr::opts_knit$set(root.dir = ("/Users/hsiaohang.tao/Dropbox/Academic/Ecoinformatics/a1_Meta_Overlap/analysis_meta_overlap/"))

```


# library
```{r}

library("lme4")
library("lmerTest")
library("tidyverse")
library("purrr")
library("broom")
library("forcats")
library("viridis")
library("boot")
library("nlme")
library("MuMIn")
library("gridExtra")
library("RColorBrewer")
library("lmtest")
library("stats")
library("cowplot")
library("synchrony")
library("dplyr")
library("stats")
library("FactoMineR")
library("factoextra")
library("readxl")
library("rfishbase")
library(boot)
library(khroma)

```


# read data & house keeping
```{r}

# read dd.co
dd.co <- read.csv("output/co.dd.csv")

dd.co <- dd.co %>% 
  mutate(species = 
           replace(species, 
                   species == "Sprattus sp",
                   "Sprattus sprattus"))

# read cpue.year
cpue.year <- read.csv('output/cpue.year.csv')

cpue.year <- cpue.year %>% 
  mutate(species = 
           replace(species, 
                   species == "Sprattus sp",
                   "Sprattus sprattus"))


# read trait data
trait.b <- read_xlsx("data/Beukhof_2019_TraitData.xlsx",1)

trait.b <- trait.b %>% 
  rename(species.orginal = species,
         species = taxon) 

name.ref <- names(dplyr::select(trait.b,contains("reference")))

trait.b <- trait.b %>% dplyr::select(-c(all_of(name.ref),
                                 family,
                                 genus,
                                 species.orginal,
                                 taxonomic.rank))



# read common.name data
common.name <- read.table("data/a_common_name_all_stocks.csv", sep = ";", header = TRUE)


# read cpue.pair
cpue.pair <- read.csv('output/cpue.pair.csv')

```



# merge trait and abundance data, create tibble [trait]
```{r}

#----- 1 North Sea (LME = 22)
#Extract species list 
ns.list <- cpue.year %>% filter(region == "North Sea") %>%
  group_by(species)  %>% tally() %>% dplyr::select(-n)

# filter to species in sp.list
ns <- inner_join(ns.list, trait.b, by = "species")
ns %>% group_by(species) %>% tally() # check num of sp
ns <- ns %>% filter(LME == 22) # all sp has ns info
ns$region <- "North Sea"

# There are two Agonus cataphractus records in trait.b,so removed one.
ns <- ns[2:nrow(ns),]

#----- 2 SWC (LME = 24)
#Extract species list 
swc.list <- cpue.year %>% filter(region == "Scottish West Coast") %>%
  group_by(species) %>% tally() %>% dplyr::select(-n)

# filter to species in sp.list
swc <- inner_join(swc.list, trait.b, by= "species")
swc %>% group_by(species) %>% tally() 
swc <- swc %>% filter(LME == 24) # all sp has info
swc$region <- "Scottish West Coast"


#-----3 Baltic Sea (LME = 23)
# Extract species list 
bs.list <- cpue.year %>% filter(region == "Baltic Sea") %>%
  group_by(species) %>% tally() %>% dplyr::select(-n)

# filter to species in sp.list
bs <- inner_join(bs.list, trait.b, by= "species")
bs %>% group_by(species) %>% tally() 
bs <- bs %>% filter(LME == 23) # all sp has info
bs$region <- "Baltic Sea"

#----- 4 Eastern Bering Sea (LME = 1)(FAO = 67 Northeast Pacific)
#Extract species list 
ebs.list <- cpue.year %>% 
  filter(region == "Eastern Bering Sea") %>%
  group_by(species) %>% 
  tally() %>% 
  dplyr::select(-n)

# filter to species in sp.list
ebs <- inner_join(ebs.list, trait.b, by= "species")
ebs %>% group_by(species) %>% tally()
ebs <- ebs %>% filter(LME == 1 & FAO == 67) # all sp has info
ebs$region <- "Eastern Bering Sea"

#----- Join all species of each region together
trait <- rbind(ns, swc, bs, ebs)

#----- remove unnessary columns
trait <- trait %>% dplyr::select(-c(LME,FAO))

```

# add common name to [trait] & [cpue.year]
```{r}

# add common name 
trait <- left_join(trait, common.name, by = c("species", "region"))

trait <- trait %>% relocate(common_name, .after = region)

cpue.year <- left_join(cpue.year, common.name, by = c("species", "region"))

cpue.year <- cpue.year %>% relocate(common_name, .after = region)

```


# [Table S1] create population trait table 

```{r}

(trait.table <- trait %>% 
  dplyr::select(common_name,species, region, growth.coefficient,
         age.maturity,age.max,length.infinity,
         length.max,fecundity) %>% 
  arrange(growth.coefficient) %>% 
  mutate(growth.coefficient = round(growth.coefficient, 2),
         age.maturity = round(age.maturity, 2),
         age.max = round(age.max, 0),
         length.infinity = round(length.infinity,0),
         length.max = round(length.max, 0)
         #,fecundity= round(fecundity / 1000, 2)
                           ))
  
write_excel_csv(trait.table,"output/trait.table.csv")

```


# add alpha from <dd.co> to <cpue.pair> as a new column

```{r}

dd.co.rename <- dd.co %>% 
  dplyr::select(region, species, year,
                size_group_1,size_group_2,
                alpha_mle) %>% 
  rename(pair_id_x = size_group_1,
         pair_id_y = size_group_2) %>% 
  mutate_at('year', as.numeric)

cpue.pair <- cpue.pair %>% 
  separate(yearqua, into = c("year", "qua")) %>%
  dplyr::select(-qua) %>%
  mutate_at('year', as.numeric)

cpue.pair <- left_join(cpue.pair, 
                              dd.co.rename,
           by = c('region','species','year',
                  'pair_id_x','pair_id_y'))

cpue.pair

```


# housekeeping <cpue.pair>

```{r}

# create [pair] from [cpue.pair]
pair <- cpue.pair %>% 
  mutate_at('pair_id', as.numeric) %>% 
  rename('cpue_x' = 'cpue_length_yearqua_x',
         'cpue_y' = 'cpue_length_yearqua_y',
         'so' = 'area_overlap')

# add group pair name
pair <- pair %>% 
  mutate_at('pair_id',as.factor) %>% 
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))

# add common name 
pair <- pair %>% 
  mutate(species = 
           replace(species, 
                   species == "Sprattus sp",
                   "Sprattus sprattus"))

pair <- left_join(pair, common.name, by = c("species", "region"))


# calculate mean cpue of x and y
pair <- pair %>% 
  mutate(mean_cpue = (cpue_x + cpue_y)/2,
         log_mean_cpue = log(mean_cpue)) 
  
  
# prepare data for logit transformation 
pair <- pair %>% 
  mutate(so = replace(so, so == 1, 0.9999),
         so = replace(so, so == 0, 0.0001),
         logit_so = logit(so)) 

# scale year
fun_scale <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

pair <- pair %>% 
  mutate_at("year", as.numeric) %>%
  mutate(scale_year = fun_scale(year)) 

```



# ----- Hypothesis I
Test if slower-growth fish has lower co-occurrence by running pair-specific regression CO ~ maximum length

## create [cpue.pair.k], [cpue.pair.k.mean]

```{r}

#1 create [trait.k] with only k as a life history trait
trait.k <- trait %>% 
  dplyr::select(region, common_name,length.max)

#2 add common name to cpue.pair 
cpue.pair.k <- left_join(cpue.pair, common.name, 
                         by = c("region","species"))

#3 add k to [cpue.pair.k]
cpue.pair.k <- left_join(cpue.pair.k, trait.k, 
          by = c("region","common_name") )

# take mean area overlap for each pair across year
cpue.pair.k.mean <- cpue.pair.k %>% 
  group_by(region, common_name, pair_id) %>% 
  mutate(mean_alpha_all_year = 
              mean(alpha_mle),
         sd_alpha_all_year= 
           sd(alpha_mle))

```


##model to extract slope store in [output.so.k]
```{r}

# nest by pair_id, region
cpue.pair.k.mean.nest  <- cpue.pair.k.mean %>% 
  dplyr::select(pair_id,region, common_name, 
         length.max, 
         mean_alpha_all_year) %>%
  group_by(pair_id, region) %>% 
  nest() %>% 
  filter(region != 'Baltic Sea')

# create functions to run regression
fun_trait_model_so <- function(df)gls(mean_alpha_all_year ~ 
    length.max, 
    na.action = na.omit, data = df)

# create function to extract output
fun_trait_extract_sep_so <-function(data){

output.trait_so <- data %>% 
  mutate(., slope = map_dbl(model, slope_fun),
                intercept = map_dbl(model, intercept_fun),
            lci = map_dbl(model, ci_lower_fun),
            uci = map_dbl(model, ci_upper_fun),
            p = map_dbl(model, p_value_fun)) %>%
  dplyr::select(pair_id,region, intercept, slope,lci, uci, p)} 

# apply two functions

output.so.k <- cpue.pair.k.mean.nest %>% 
  mutate(model = purrr::map(data, fun_trait_model_so))

output.so.k <- fun_trait_extract_sep_so(output.so.k)


# assign p-sig
output.so.k <- output.so.k %>% 
  mutate(p_sig = replace(p, p<0.05,"sig")) %>% 
  mutate(p_sig = replace(p_sig, p > 0.05, "sig-non"))%>% 
  relocate(p_sig,.after = "p")


cpue.pair.k.mean.nest
```


## [Fig 2a] plot alpha~maximum length relationship by pair
output 3 x 7
```{r}

cpue.pair.k.mean <- cpue.pair.k.mean %>%
  mutate(region_reorder = case_match(region,
      "North Sea" ~ "1_North Sea",
      "Scottish West Coast"~ "2_Scottish West Coast",
      "Eastern Bering Sea" ~ "3_Eastern Bering Sea")) %>%
  mutate_at('pair_id', as.factor) %>%
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))


# plot all pairs
cpue.pair.k.mean %>%
ggplot(.,aes(x = length.max,
                 y = mean_alpha_all_year))+
  geom_point(aes(color = group_pair), size = 0.1)+
  geom_smooth(method = 'lm', se= F, aes(color = group_pair),
              size = 0.4)+
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank())+
  facet_wrap(vars(region), ncol = 3)+
  ylab("Alpha MLE")+xlab('Maximum body size (cm)')+
  facet_wrap(vars(region_reorder), ncol = 4,
             labeller = labeller(region_reorder = 
    c("1_North Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      "3_Eastern Bering Sea" = "Eastern Bering Sea")))+
  #ylim(-3, 5)+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))


```


## [Fig 2b] forest plot
output 4 x 6
```{r}

# add add <region_reorder> & <group_pair> to [output.so.k]
output.so.k <- output.so.k %>% 
  mutate(region_reorder = case_match(region,
      "North Sea" ~ "1_North_Sea",
      "Scottish West Coast"~ "2_Scottish West Coast",
      "Eastern Bering Sea" ~ "3_Eastern Bering Sea")) %>%
  mutate_at('pair_id', as.factor) %>% 
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))

# forest plot
output.so.k %>% 
  ggplot(., 
         aes(y = group_pair, x = slope))+
  geom_point(shape = 16, aes(color = group_pair),
             size = 2)+
  geom_linerange(aes(xmin=lci, xmax=uci,
                     linetype = p_sig),
                 size = 0.5)+
  facet_wrap(vars(region_reorder), ncol = 4,
             labeller = labeller(region_reorder = 
    c("1_North_Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      "3_Eastern Bering Sea" = "Eastern Bering Sea")))+
  theme_bw()+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, size=0.3, linetype = 'solid')+
  xlab('slope and CI of Alpha MLE ~ Maximum body size')+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))+
  scale_y_discrete(limits=rev)+
  scale_x_continuous(breaks = seq(-0.02, 0, by = 0.01))


```



#-----Hypothesis II

examine CO~abundance for each population

#1. prepare [pair.nest] for applying functions

```{r}

pair.nest <- pair %>% 
  ungroup() %>% 
  dplyr::select(pair_id, region, common_name, 
         scale_year, 
         log_mean_cpue,
         logit_so,
         alpha_mle) %>%
  group_by(pair_id, region, common_name) %>% 
  nest()

```


#2. create function to model CO~abundance for each pair, store in [mod.data.pair]

```{r}

# create the function
fun_pair_slope <- function(df)gls(alpha_mle~ 
    log_mean_cpue + scale_year, 
  na.action = na.omit, data = df)

# run the function
pair.nest <- pair.nest %>% dplyr::filter(region != 'Baltic Sea')

mod.data.pair <- pair.nest %>% 
  mutate(model = purrr::map(data, fun_pair_slope))


```


#3. create function to extract slope & CI and run, from [mod.data.pair], store in [output.pair]

```{r}

# functions to extract slope & CI from model
slope_fun <- function(mod)coefficients(mod)[[2]]
ci_lower_fun <- function(mod) confint(mod)[2,1]
ci_upper_fun <- function(mod) confint(mod)[2,2]
p_value_fun <- function(mod)summary(mod)$tTable[2,4]
resid_fun <- function(mod)resid(mod)
AIC_fun <- function(mod)AIC(mod)
intercept_fun <- function(mod)coefficients(mod)[[1]]

fun_extract_mod <-function(data){
  output <- data %>% 
  dplyr::mutate(., slope = map_dbl(model, slope_fun),
            lci = map_dbl(model, ci_lower_fun),
            uci = map_dbl(model, ci_upper_fun),
            p = map_dbl(model, p_value_fun)) %>%
  dplyr::select(pair_id, region, common_name, slope, lci, uci, p)
} 

# apply functions
output.pair <- fun_extract_mod(mod.data.pair)

```


#4 plot [Fig 3b]: beta for all populations per region 
output 6 x 9
```{r}

# prepare [output.pair] for plot
## 1 add p_sig
output.pair <- output.pair %>% 
  mutate(p_sig = replace(p, p<0.05,"sig")) %>% 
  mutate(p_sig = replace(p_sig, p > 0.05, "sig-non"))%>% 
  relocate(p_sig,.after = "p") %>% 
  mutate(region_reorder = case_match(region,
      "North Sea" ~ "1_North Sea",
      "Scottish West Coast"~ "2_Scottish West Coast",
      #'Baltic Sea' ~ '3_Baltic Sea',
      "Eastern Bering Sea" ~ "4_Eastern Bering Sea"))

## 2 add trait to reorder common_name
trait.name.k <- trait %>% 
  dplyr::select(region, common_name, length.max)

output.pair <- inner_join(output.pair, trait.name.k, 
           by = c('region','common_name'))

output.pair$common_name <-
  fct_reorder(output.pair$common_name,                                 desc(output.pair$length.max))


# add column group_pair
output.pair <- output.pair %>% 
  mutate_at('pair_id',as.factor) %>% 
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))

output.pair %>% 
ggplot(data=., 
       aes(y=common_name, x=slope)) +
  geom_linerange(aes(xmin= lci, xmax= uci, alpha = 0.5, 
                     color = group_pair,
                     linetype = p_sig),
                 position=position_jitter(w=0, h=0.25))+ 
  facet_wrap(vars(region_reorder), ncol = 4,
             labeller = labeller(region_reorder = 
    c("1_North Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      #"3_Baltic Sea" = "Baltic Sea",
      "4_Eastern Bering Sea" = "Eastern Bering Sea")))+
  geom_point(shape = 16,size = 0.01,
             aes(color = group_pair)) + 
  theme_bw()+
  guides(color = guide_legend(reverse=TRUE))+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, linetype = 'dashed', size = 0.3)+
  xlab('beta SO +- se')+ ylab('')+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))+
  #xlim(-2, 4)+
  geom_jitter(aes(color = group_pair),
              width = 0.05, height = 0.25)
 
```


#5. plot [Fig 3a]: example population for each region
output 6 x 8
```{r}

# plot function
function_plot_example <- function(data, ggtitle){
  #@20250121 y change to alpha_mle
  ggplot(data = data, 
       aes(x = log_mean_cpue, 
           y = alpha_mle)) +
  geom_point(aes(color = group_pair),
             shape = 16)+
  geom_smooth(method = "lm", formula = y~x, se = F,  
              aes(color = group_pair),
              size = 0.5)+
  theme_bw()+
  xlab("mean abundance of the pair (log transformed)") +
  ylab("co-occurrence (alpha mle)") +
  # remove background
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank())+
  ggtitle(ggtitle)+
     scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134',
           '#FF6E3A','#FFC33B',"#000000"))
  }


# extract 1 fish per region

pair.ns <- pair %>% 
  filter(region == 'North Sea' &
           common_name == 'Atlantic cod') %>% 
  mutate_at('group_pair', as.factor)

pair.swc <- pair %>% 
  filter(region == 'Scottish West Coast' &
           common_name == 'Red gurnard') %>% 
  mutate_at('group_pair', as.factor)

# pair.bal <- pair %>% 
#   filter(region == 'Baltic Sea' &
#            common_name == 'Sprat') %>% 
#   mutate_at('group_pair', as.factor)

pair.ebs <- pair %>% 
  filter(region == 'Eastern Bering Sea' &
           common_name == 'Pacific halibut') %>% 
  mutate_at('group_pair', as.factor)

a1 <- function_plot_example(pair.ns,'N.S. Atlantic cod')
a2 <- function_plot_example(pair.swc, 'S.W.C. Hooknose')
#a3 <- function_plot_example(pair.bal, 'B.S. Sprat')
a4 <- function_plot_example(pair.swc, 'E.B.S. Pacific halibut')

grid.arrange(a1, a2, a4, ncol=2)

```



# 6 ratio of positive relationships

```{r}

# number of negative & positive relationships

output.pair %>% 
  filter(slope < 0) # 165

output.pair %>% 
  filter(slope > 0) # 375


# total number of relationships: 54 populations x 10 size group pairs = 540 

165+375 # sum = 540

# ratio of positive relationship / all relationship
165/540 # 31%
375/540 # 69%


```

#7 ratio of sig over all relationships

```{r}

# ratio of signigicance out of all positive relationships
output.pair %>% 
  # 375 pairs showed slope >0
  filter(slope > 0) %>% 
  # 84 of them is significant
  filter(p_sig == 'sig')

# ratio 
84/375 # 22%
84/540 # 15% out of all pairs

# ratio of signigicance out of all negative relationships
output.pair %>% 
  # 165 pairs showed slope <0
  filter(slope < 0) %>% 
  # 17 of them is significant
  filter(p_sig == 'sig')

# ratio 
17/165 # 10%
17/540 # 3% out of all pairs



```


#8 check proportion sig with population

```{r}

# number of population with more than 50% of size group pairs with significant positive relationship

output.pair %>% 
  group_by(p_sig, region, common_name, slope) %>%
  filter(slope > 0 & p < 0.05) %>% 
  group_by(region, common_name) %>% 
  tally() %>% 
  filter(n>4)
  

```

# 9 check by region

result:
number of populations in each region: 2,8,27,19
number of populations with more than 50% group has sig p: 0,3,21,9

```{r}

# number of populations in each region: 2,8,27,19
output.pair %>% filter(pair_id == 1) %>% group_by(region) %>% tally
  


```

# 10 check proportion sig with pair id
Result
pair_id: 61%-79% are significant
prop_sig from high to low: 5,1,8,10,9,6,2,4,7,3
```{r}

output.pair %>% 
  group_by(pair_id, group_pair,p_sig) %>%
  tally() %>% 
  rename(number = n) %>% 
  pivot_wider(names_from = p_sig, values_from = number) %>% 
  rename(sig_non = 4) %>% 
  mutate(sum = sig+ sig_non) %>% 
  mutate(prop_sig = sig / sum) %>% 
  arrange(desc(prop_sig))
  
```


#-----Hypothesis III

between populations: examine betaCO ~ life history 

# prepare data
join [output.pair] with [trait] to :
1. prepare [trait.pair] to make a graph, and 
2. prepare [trait.pair.nest] to model beta ~ length.max
```{r}

#1 select 4 additional life history traits, create [trait.4]
trait.4 <- trait %>% ungroup() %>% 
  dplyr::select(region, common_name,
                age.maturity, length.infinity,
                age.max, growth.coefficient, length.max)

#2 merge [trait.4] and [output.pair]
trait.pair <- left_join(output.pair, trait.4, by = c("region","common_name"))

#2 group trait.pair, create [trait.pair.sep]
trait.pair.nest <- trait.pair %>% 
  pivot_longer(c(growth.coefficient, age.maturity,
                 length.infinity,length.max,
                 age.max),
               names_to = "trait.name",
               values_to = "trait.value")

#3 nest by pair_id, region, response, trait.name
trait.pair.nest  <- trait.pair.nest  %>% 
  dplyr::select(pair_id,region, common_name, 
         slope, trait.name, trait.value) %>%
  group_by(pair_id, region, trait.name) %>% 
  nest()

```


# model beta ~ life history, create [output.pair.beta]
```{r}

# function: model beta
fun_trait_model <- function(df)gls(slope ~ 
    trait.value, na.action = na.omit, data = df)

# function: extract output
fun_trait_extract_sep <-function(data){

output.trait <- data %>% 
  mutate(., slope = map_dbl(model, slope_fun),
                intercept = map_dbl(model, intercept_fun),
            lci = map_dbl(model, ci_lower_fun),
            uci = map_dbl(model, ci_upper_fun),
            p = map_dbl(model, p_value_fun)) %>%
  dplyr::select(pair_id,region, trait.name, slope, intercept, lci, uci, p)

} 

# apply function
output.pair.beta <- trait.pair.nest %>% 
  mutate(model = purrr::map(data, fun_trait_model))

output.pair.beta <- fun_trait_extract_sep(output.pair.beta)

output.pair.beta <- output.pair.beta %>% arrange(trait.name)

#5 assign p-sig
output.pair.beta <- output.pair.beta %>% 
  mutate(p_sig = replace(p, p<0.05,"sig")) %>% 
  mutate(p_sig = replace(p_sig, p > 0.05, "sig-non"))%>% 
  relocate(p_sig,.after = "p")
  
```


# calculate trait range, expected y, save [output.pair.beta]
This is to prepare for xyplot of beta ~ life history
Steps:
1. use [trait.pair.nest] trait values to generate max and min trait <trait.max> <trait.min> , store in [x.max.min]
2. join [x.max.min] to [output.pair.beta]
3. in [output.pair.beta] calculate <trait.yend> using intercept, slope, and trait.max

```{r}

# extract x range (it is region-specific depending on which species included in each region)

#1 data
trait.pair.nest
  
#2 function to extract x range
fun_x <- function(df){
  df %>%
    summarise(trait.min = min(trait.value),
              trait.max = max(trait.value))}

#3 apply function
x.max.min <- trait.pair.nest %>%
  ungroup() %>% 
  group_by(region) %>% 
  mutate(output = purrr::map(data, fun_x))

#4 arrange output to create a x range table
x.max.min  <- x.max.min  %>% 
  ungroup() %>% 
  dplyr::select(pair_id, region, trait.name, output) %>% 
  unnest(cols = c(output))


# add x range to out.trait.sep
output.pair.beta <- 
  left_join(output.pair.beta, x.max.min, 
            by = c('pair_id',"region", "trait.name"))

# calculate y range (ymin = intercept so not need to calculate. Only calculate yend)
output.pair.beta <- output.pair.beta  %>% 
  mutate(trait.yend = intercept + slope * trait.max)


```

# re-order region, rename pair for [trait.pair] [output.pair.beta]
```{r}

# check pair id of x & y
cpue.pair %>% 
  group_by(pair_id) %>% 
  summarise(x = mean(pair_id_x),
            y = mean(pair_id_y))

# for trait.pair
trait.pair <- 
  trait.pair %>%
  mutate(region_reorder = case_match(region,
      "North Sea" ~ "1_North_Sea",
      "Scottish West Coast"~ "2_Scottish West Coast",
      "Eastern Bering Sea" ~ "3_Eastern Bering Sea")) %>%
  mutate_at('pair_id', as.factor) %>% 
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))

# for output.pair.beta
output.pair.beta <- 
  output.pair.beta %>%
  mutate(region_reorder = case_match(region,
      "North Sea" ~ "1_North_Sea",
      "Scottish West Coast"~ "2_Scottish West Coast",
      "Eastern Bering Sea" ~ "3_Eastern Bering Sea")) %>%
  mutate_at('pair_id', as.factor) %>% 
  mutate(group_pair = case_match(pair_id,
                              '1' ~ '1-2',
                              '2' ~ '1-3',
                              '3' ~ '1-4',
                              '4' ~ '1-5',
                              '5' ~ '2-3',
                              '6' ~ '2-4',
                              '7' ~ '2-5',
                              '8' ~ '3-4',
                              '9' ~ '3-5',
                              '10' ~ '4-5'))


```


# plot [Fig 4a]: CO ~ max length in xy plot

```{r}


# using [output.pair.beta] to: 
## 1 create [slope.pair], for adding to ggplot2 to assign trait.min, trait.max, intercept, trait.yend

slope.pair <- output.pair.beta %>% 
    filter(trait.name == "length.max") %>% 
  mutate_at('group_pair', as.factor)

## 2 create [line.type], for assigned line types 
slope.pair.line <- ifelse(slope.pair$p_sig =="sig", 
                               "solid", "dashed")


# plot all pairs
trait.pair %>%
ggplot(.,aes(x = length.max,
                 y = slope, ymin =lci, ymax = uci))+
  geom_pointrange(aes(color = group_pair),
                  shape = 16, size = 0.3) +
  # geom_smooth(method = 'lm', aes(color = pair_id),
  #             se = FALSE, size = 0.5)+
  geom_segment(data = slope.pair,
      aes(x = trait.min, xend = trait.max, 
          y = intercept, yend = trait.yend,
          color = group_pair), 
      size = 0.4,
      lty = slope.pair.line)+
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    legend.position="none")+
  facet_wrap(vars(region_reorder), ncol = 3,
             labeller = labeller(region_reorder = 
    c("1_North_Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      "3_Eastern Bering Sea" = "Eastern Bering Sea")))+
  ylab("pair-specific beta")+
  ylim(-3, 5)+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))+
  xlab('Maximum body size (cm)')+
  ylab('betaCO')




# another color example
colorBlindBlack8  <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")



```


# plot [Fig 4b]: forest plot beta CO ~ max length

```{r}
  
output.pair.beta %>% 
  mutate_at('pair_id', as.factor) %>%
  filter(trait.name == 'length.max') %>% 
  ggplot(., 
         aes(y = group_pair))+
  geom_point(aes(x = slope,color = pair_id), 
             shape = 16, size = 3)+
  geom_linerange(aes(xmin=lci, xmax=uci,
                     linetype = p_sig, alpha = 0.5),
                 size = 0.8)+
  facet_wrap(vars(region_reorder),
             labeller = labeller(region_reorder = 
    c("1_North_Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      "3_Eastern Bering Sea" = "Eastern Bering Sea")))+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))+
  theme_bw()+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, size=0.3, linetype = 'solid')+
  xlab('betaCO-maximum body size (slope +- CI)')+
  ylab('')+
  scale_y_discrete(limits=rev)


```


# plot [Fig S1] plot other life history traits

```{r}

plot.life.history <- function(life.history, title){
  output.pair.beta %>% 
  mutate_at('pair_id', as.factor) %>%
  filter(trait.name == life.history) %>% 
  ggplot(., 
         aes(y = group_pair))+
  geom_point(aes(x = slope,color = pair_id), 
             shape = 16, size = 3)+
  geom_linerange(aes(xmin=lci, xmax=uci,
                     linetype = p_sig, alpha = 0.5),
                 size = 0.8)+
  facet_wrap(vars(region_reorder),
             labeller = labeller(region_reorder =
    c("1_North_Sea" = "North Sea",
      "2_Scottish West Coast" = "Scottish West Coast",
      "3_Eastern Bering Sea" = "Eastern Bering Sea")))+
  scale_color_manual(values = c('#9F0162','#009F81','#8400CD','#008DF9',
           '#00C2F9','#FFB2FD','#E20134','#FF6E3A','#FFC33B',"#000000"))+
  theme_bw()+
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  scale_y_discrete(limits=rev)+
  theme(legend.position = "none")+
  geom_vline(xintercept = 0, size=0.3, linetype = 'solid')+
  xlab('')+
  ggtitle(title)
}

# traits: age.maturity,age.max,length.infinity,length.max, fecundity

p1 <- plot.life.history('length.infinity','Infinite body size')
p2 <- plot.life.history('age.max','Maximum age')
p3 <- plot.life.history('age.maturity','age at maturity')
p4 <- plot.life.history('growth.coefficient','growth coefficient k')

grid.arrange(p1, p2, p3,p4, ncol=2)
```




#-------statistics
# summary: list of stocks
```{r}


# number of stocks per region
# 27 NS, SWC 19, EBS 8 
cpue.pair %>% group_by(region) %>% 
  summarise(number_species = n_distinct(species))

# total number of stocks: 54
cpue.pair %>% group_by(region, species) %>% 
  tally()

# total number of species: 38
cpue.pair %>% group_by(species) %>% 
  summarise(number_species = n_distinct(species))

```



