---
title: "0_format_raw_data"
output: html_document
editor_options: 
  chunk_output_type: inline
---
#Set working directory
```{r setup}

# knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file("/Users/jhih-shengliu/Desktop/Fish_overlap/NS-IBTS/NS-IBTS_JSL/"))


knitr::opts_knit$set(root.dir = ("/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/"))


```


#library
```{r}
library("tidyverse")
library("tidyr")

```

#------------North Sea--------------------
# Call North Sea raw data to check column name
```{r}
 
ns.fish.raw <- read.csv("data/raw_data/NS_1977-2021Q1.csv") 
names(ns.fish.raw)

ns <- ns.fish.raw %>%
  select(-X.1,-X,-Survey,-Area,-AphiaID,-DateofCalculation) %>%
  rename(year=Year,
         quarter=Quarter,
         subarea=SubArea,
         species=Species,
         length=LngtClass,
         cpue_number_per_hour=CPUE_number_per_hour,
         region=Region)

ns <- ns %>% 
  relocate("region",.before = "year") %>% 
  mutate_at("species", as.character)

```

# Save NS 
```{r}
write.csv(ns,file="data/raw_data/ns_raw_formatted.csv")
```

#------------Scotish West Coast--------------------
# Call Scotish West Coast raw data to check column name
```{r}
 
swc.fish.raw <- read.csv("data/raw_data/SWC_all.csv")
swc.fish.raw$region <-"Scottish West Coast"
names(swc.fish.raw)

swc <- swc.fish.raw %>%
  select(-Survey,-Area,-AphiaID,-DateofCalculation) %>%
  rename(year=Year,
         quarter=Quarter,
         subarea=SubArea,
         species=Species,
         length=LngtClass,
         cpue_number_per_hour=CPUE_number_per_hour)

swc <- swc %>%
  relocate("region",.before = "year") %>% 
  mutate_at("species", as.character)

```

# Save SWC 
```{r}
write.csv(swc,file="data/raw_data/swc_raw_formatted.csv")
```


#------------Baltic--------------------

# Baltic: herring & sprat
```{r}
getwd()

bal.herring <- read.csv("data/raw_data/baltic_sea_raw/bal_herring_data_raw_mich.csv")

bal.herring <- bal.herring %>%
 pivot_longer(cols = starts_with("X"),
              names_to = "length",
              names_prefix = "X",
              values_to = "CPUE") %>%
 mutate_at("length", as.numeric) %>%
 mutate(scientific_name = "Clupea harengus") %>%
 rename("year"="Year")


# sprat
bal.sprat <- read.csv("data/raw_data/baltic_sea_raw/bal_sprat_data_raw_mich.csv")


bal.sprat <- bal.sprat %>%
 pivot_longer(cols = starts_with("X"),
              names_to = "length",
              names_prefix = "X",
              values_to = "CPUE") %>%
 mutate_at("length", as.numeric) %>%
 mutate(scientific_name = "Sprattus sp")%>%
 rename("year"="Year")


# bind herring + sprat
bal <- bind_rows(bal.herring, bal.sprat)
names(bal)

bal <- bal %>%
 rename("cpue_number_per_hour"="CPUE",
        "subarea" = "Rec") %>%
 mutate(region = "baltic sea") %>%
 select(-SD)

# There is no quarterly data for bal, so set quarter = 0
bal$quarter = 0
bal$region <- "Baltic Sea"
bal <- bal %>%
 rename(species=scientific_name) %>%
 select("region","year", "quarter","subarea",
        "species","length","cpue_number_per_hour") %>%
 mutate_at("subarea", as.character)

```



# Save Bal (herring + spart)
```{r}

write.csv(bal,file="data/raw_data/bal_raw_formatted.csv")

```


#--------- Eastern Bering Sea-------

# Read data
```{r}

# group-read csvs of ebs data

Path <- "/Users/sha/Documents/Ecoinformatics/a1_Meta_Overlap/Analysis/data/raw_data/ebs_nbs_paul_raw/"

filenames <- list.files(path = Path,
                       pattern = "ebs.*csv")

filepath <- paste0(Path, filenames)

ebs <- lapply(filepath, read.csv)

ebs <- ebs %>% bind_rows() %>% 
  select(LATITUDE, LONGITUDE,
               SPECIES_NAME, COMMON_NAME,
               LENGTH,YEAR,CPUE_LENGTH) %>% 
  add_column(region ="EBS",.before = "LATITUDE") 
  

names(ebs) <- c("region","lat","lon","scientific_name",
                "species", "length","year",
                "cpue_number_per_hour")

```

# generate list of scientific_name & common name 
```{r}

ebs.species.list <- ebs %>% 
  group_by(scientific_name, species) %>% 
  tally() %>% 
  select(scientific_name, species)

# # save list
# write.csv(ebs.species.list,
#           file="data/Raw_Data/ebs_species_list.csv")

```


# Convert length from cm to mm
```{r}

ebs <- ebs %>% mutate(length_mm = length * 10) %>% 
  select(-length) %>% 
  rename(length = length_mm)

  
```


# Grid the data into 1 x 0.5 
```{r}

# split lon
#1 separate interger and decimal of longitude coordinates
lon <- do.call(rbind, strsplit(as.character(ebs$lon),"\\."))
#2 add interger and decimal columns into ebs tibble
ebs <- bind_cols (ebs, as_tibble(lon))
#3 rename interger and decimal columns in ebs
ebs <- ebs %>% rename(lon_int = V1, lon_dec = V2)

# split lat
#1 round latitudinal coordinates
ebs <- ebs %>% mutate(across(lat, round, 2))

#2 separate interger and decimal of longitude coordinates
lat <- do.call(rbind, strsplit(as.character(ebs$lat),"\\."))

#3 add interger and decimal columns into ebs tibble
ebs <- bind_cols (ebs, as_tibble(lat))

#4 rename interger and decimal columns in ebs
ebs <- ebs %>% rename(lat_int = V1, lat_dec = V2)

# 5 assign lat_dec > 50: 1, lat_dec <50:2
ebs <- ebs %>% mutate(lat_dec_50 =
                 replace(lat_dec, lat_dec >50, "1")) %>% mutate(lat_dec_50 = replace(lat_dec_50, lat_dec < 50, "2"))

#6 bind interger and decimal (either 1 or 2)
ebs <- ebs %>% 
  mutate(lat_new=paste(lat_int,lat_dec_50,sep=".")) %>% 
  mutate_at("lat_new", as.numeric)

```


# aggregate CPUE based on 1 x 0.5 grid
Taking the mean of cpue_number_per_hour of all survey points within a grid 
```{r}

ebs.grid <- ebs %>% 
  ungroup() %>% 
  group_by(region, scientific_name, length,
           species, year, lon_int, lat_new) %>% 
  summarise(cpue_number_per_hour = mean(cpue_number_per_hour),
            .groups = 'drop') 

# name the grids (there are 184 grids)
subarea <- ebs.grid %>% group_by(lon_int, lat_new) %>% tally() %>% select(lon_int, lat_new) 

subarea$subarea = c(1:184)

ebs <- full_join(ebs.grid, subarea, by = c("lon_int", "lat_new"))

names(ebs)

ebs <- ebs %>% ungroup() %>% dplyr::select(-c("lat_new","lon_int"))

# check survey period (1982-2019)
ebs %>% ungroup() %>% 
  group_by(year) %>% 
  tally()

```

# Edit column name & quarter data
```{r}
ebs <- ebs %>% select(-species)
names(ebs)

#------- EBS
ebs$quarter = 0
ebs$region <- "Eastern Bering Sea"

ebs <- ebs %>%
  rename(species=scientific_name) %>%
  select("region","year", "quarter","subarea",
         "species","length","cpue_number_per_hour") %>%
  mutate_at("subarea", as.character)

```


# check rare species
```{r}

ebs %>% 
  # take sum of cpue from all lengths of all years
  group_by(species) %>% 
  summarise(cpue_all_length_year = 
              sum(cpue_number_per_hour)) %>% 
  arrange(cpue_all_length_year)

```

# Save EBS
```{r}
write.csv(ebs,file="data/Raw_Data/ebs_raw_formatted.csv")

```



